---
title: "UKB_ALC_BMD"
author: "Xinpeng"
date: '2024-06-17'
output:
  word_document: default
  html_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = '//store/department/im/share/personal/059909/PHD/Bone & Alcohol/UKBB_data/UKB_alc_and_bmd/UKB_ALC_BMD/')

library(dplyr)
library(ggplot2)
library(ggpubr)
library(AER)
library(tableone)
library(broom)
library(readxl)
library(readr)
library(tidyverse)
library(forestplot)
```



# creating datasets
```{r}

ukb_alc_bmd <- read_csv("//store/department/im/share/personal/059909/PHD/Bone & Alcohol/UKBB_data/UKB_alc_and_bmd/UKB_ALC_BMD/dataset_alc_bmd.csv")
ukb_alc_bmd <- ukb_alc_bmd[,-1]
# 38715 participants with BMD measurement
```

```{r}
ukb_alc_bmd$eth <- ukb_alc_bmd$`21000-0.0`
ukb_alc_bmd$geneticEthnic <- ukb_alc_bmd$`22006-0.0`
ukb_alc_bmd$sex <- ukb_alc_bmd$`31-0.0`
ukb_alc_bmd$age <- ukb_alc_bmd$`21003-2.0`
ukb_alc_bmd$BMI <- ukb_alc_bmd$`21001-2.0`
ukb_alc_bmd$income <- ukb_alc_bmd$`738-2.0`
ukb_alc_bmd$education <- ukb_alc_bmd$`6138-2.0`
ukb_alc_bmd$smkstatus <- ukb_alc_bmd$`20116-2.0`
ukb_alc_bmd$alcFrq <- ukb_alc_bmd$`1558-2.0`
ukb_alc_bmd$beer <- ukb_alc_bmd$`1588-2.0`
ukb_alc_bmd$spirits <- ukb_alc_bmd$`1598-2.0`
ukb_alc_bmd$fortified <- ukb_alc_bmd$`1608-2.0`
ukb_alc_bmd$champagne <- ukb_alc_bmd$`1578-2.0`
ukb_alc_bmd$redwine <- ukb_alc_bmd$`1568-2.0`
ukb_alc_bmd$otheralc <- ukb_alc_bmd$`5364-2.0`
ukb_alc_bmd$fnbmd <- ukb_alc_bmd$`23208-2.0`
ukb_alc_bmd$lsbmd <- ukb_alc_bmd$`23204-2.0`
ukb_alc_bmd$tobmd <- ukb_alc_bmd$`23236-2.0`
ukb_alc_bmd$beer_month <- ukb_alc_bmd$`4429-2.0`
ukb_alc_bmd$spirits_month <- ukb_alc_bmd$`4440-2.0`
ukb_alc_bmd$fortified_month <- ukb_alc_bmd$`4451-2.0`
ukb_alc_bmd$champagne_month <- ukb_alc_bmd$`4418-2.0`
ukb_alc_bmd$redwine_month <- ukb_alc_bmd$`4407-2.0`
ukb_alc_bmd$otheralc_month <- ukb_alc_bmd$`4462-2.0`


ukb_alc_bmd <- ukb_alc_bmd %>%
  mutate_all(~ ifelse(. %in% c(-1,-3,-7), NA, .))
```

```{r}
df <- ukb_alc_bmd[,c(1,32:56)]
df <- df[!is.na(df$alcFrq),] #38425 participants with alcohol frequency

# calculate alcohol intake
df$alctot <- df$beer*20+(df$spirits+df$fortified+df$champagne+df$redwine)*10+df$otheralc*0
df$alctot_month <- df$beer_month*20+(df$spirits_month+df$fortified_month+df$champagne_month+df$redwine_month)*10+df$otheralc_month*0

df$alctot_1 <- df$alctot/7
df$alctot_month_1 <- df$alctot_month/30.4375
df$alctot0 <- ifelse(!is.na(df$alctot_1),df$alctot_1,df$alctot_month_1)
df$alctot2 <- ifelse(df$alcFrq==6,0,df$alctot0)

#df2 <- df[!is.na(df$alctot0),]
#df2$alctot_1 <- df2$alctot/7
#df2$alctot_month_1 <- df2$alctot_month/30.4375
#df2$alctot0 <- ifelse(!is.na(df2$alctot_1),df2$alctot_1,df2$alctot_month_1)
df$alctot2 <- ifelse(df$alcFrq==6,0,df$alctot0)

df4 <- df[!is.na(df$alctot2),] # filter out participants without alcohol consumption 38096 left

df4 <- df4 %>%
  mutate(eth_all = case_when(
    eth %in% c(1, 1001, 1002, 1003) ~ 1,
    eth %in% c(2, 2001, 2002, 2003, 2004) ~ 2,
    eth %in% c(3, 3001, 3002, 3003, 3004) ~ 3,
    eth %in% c(4, 4001, 4002, 4003) ~ 4,
    eth %in% c(5, 6) ~ 5
  ))

df5 <- df4[!is.na(df4$geneticEthnic),] # 32323 participants left
```


```{r}
# filtering out related participants randomly
relatedness1 <- read.table("//store/department/im/share/projects/Fernando_Group/Alcohol_BMD_Xinpeng/UK Biobank/datasets/ukb_rel_a67864_s488161.txt",header = T)

relatedness1 <- relatedness1 %>%
  mutate(relatedness = if_else(Kinship >= 0.0884, 1,0))
table(relatedness1$relatedness)
relatedness1 <- relatedness1 %>%
  rename(eid = ID1)
relatedness_0.0884 <- relatedness1[relatedness1$relatedness==1,] 

relatedness1 <- relatedness1 %>%
  mutate(relatedness_0.0442 = if_else(Kinship >= 0.0442, 1, 0))
table(relatedness1$relatedness_0.0442)
relatedness_0.0442 <- relatedness1[relatedness1$relatedness_0.0442 == 1, ]

set.seed(123)
unique_data <- relatedness_0.0442 %>%
  group_by(eid) %>%
  sample_n(1) %>%
  ungroup()
print(unique_data)
write.csv(unique_data, "relatedness_0.0442_unique.csv")

df5 <- left_join(df5, unique_data, by = "eid")
table(df5$relatedness,useNA = "always")
df6 <- df5 %>%
  filter(is.na(relatedness) | relatedness != 1)
dim(df6)
df6$PA_all_hrs <- df6$PA_all/60
df6 <- df6 %>%
  mutate(edu_recode = case_when(
                     education %in% c(3,4) ~ 1,
                     education == 2 ~ 2,
                     education == 6 ~ 3,
                     education %in% c(1,5) ~ 4))
```

# Basic characteristics
```{r}
myVars <- c("age","sex","BMI","geneticEthnic","education","income","smkstatus","alcFrq","PA_all","beer","spirits","fortified","champagne","redwine","otheralc","alctot2","lsbmd","fnbmd","tobmd","PA_all_hrs","edu_recode")
catVars <- c("sex","education","income","smkstatus","alcFrq","edu_recode")
nonormalVars <- c("PA_all","alctot2","PA_all_hrs")

tab2 <- CreateTableOne(vars = myVars,data = df6,factorVars = catVars,strata = "sex")
tab2Mat <- print(tab2,nonnormal = nonormalVars,formatOptions = list(big.mark=","))
write.csv(tab2Mat, "DescTab_UKBB.csv")
summary(tab2)
```
```{r}
tab3 <- CreateTableOne(vars = myVars, data = df6, factorVars = catVars, strata = "alc_group")
tab3Mat <- print(tab3, nonnormal = nonormalVars, formatOptions = list(big.mark=","), showAllLevels = TRUE)
write.csv(tab3Mat, "DescTab_UKBB_withoutRelatedness.csv")
```

 Histogram of alcohol consumption
```{r}
hist(df6$alctot2, main = "Histogram of alcohol intake", xlab = "log(alcohol/10+1)")
hist(log(df6$alctot2/10+1), main = "Histogram of log-transformed alcohol intake", xlab = "log(alcohol/10+1)")
```

# Linear model - general population
```{r}
# LSBMD
df6$sex <- as.factor(df6$sex)
df6$income <- as.factor(df6$income)
df6$education <- as.factor(df6$education)
df6$smkstatus <- as.factor(df6$smkstatus)
df6$eth_all <- as.factor(df6$eth_all)
df6 <- left_join(df6,ukb_lstbs_origin, by="eid")
df6 <- df6 %>% 
  group_by(sex) %>% 
  mutate(std_lsbmd = as.vector(scale(lsbmd)),
         std_fnbmd = as.vector(scale(fnbmd)),
         std_tobmd = as.vector(scale(tobmd)),
         std_lstbs = as.vector(scale(LSTBS))) %>% 
  ungroup()



lsmod1 <- lm(std_lsbmd~log(alctot2/10+1)+age+sex,data = df6)
lsmod2 <- lm(std_lsbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all,data = df6)
lsmod3 <- lm(std_lsbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
lsmod3_interaction <- lm(std_lsbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI+log(alctot2/10 + 1):age,data = df6)

tidy(lsmod1, conf.int = T)
tidy(lsmod2, conf.int = T)
tidy(lsmod3, conf.int = T)
tidy(lsmod3_interaction, conf.int=T)
```

```{r}
qqnorm(lsmod3$residuals, main = "Normal Q-Q Plot for residuals", xlab = "Residuals")
```

```{r}
# NKBMD
fnmod1 <- lm(std_fnbmd~log(alctot2/10+1)+age+sex,data = df6)
fnmod2 <- lm(std_fnbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all,data = df6)
fnmod3 <- lm(std_fnbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)

tidy(fnmod1, conf.int = T)
tidy(fnmod2,conf.int = T)
tidy(fnmod3,conf.int = T)
```
```{r}
qqnorm(fnmod3$residuals, main = "Normal Q-Q Plot for residuals", xlab = "Residuals")
```

```{r}
# TOBMD
tomod1 <- lm(std_tobmd~log(alctot2/10+1)+age+sex,data = df6)
tomod2 <- lm(scale(tobmd)~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all,data = df6)
tomod3 <- lm(scale(tobmd)~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)

tidy(tomod1,conf.int = T)
tidy(tomod2,conf.int = T)
tidy(tomod3,conf.int = T)
```

```{r}
qqnorm(tomod3$residuals, main = "Normal Q-Q Plot for residuals", xlab = "Residuals")
```

```{r}
# LSTBS
tbsmod1 <- lm(std_lstbs~log(alctot2/10+1)+age+sex,data = df6)
tbsmod2 <- lm(std_lstbs~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all,data = df6)
tbsmod3 <- lm(std_lstbs~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)

tidy(tbsmod1,conf.int = T)
tidy(tbsmod2,conf.int = T)
tidy(tbsmod3,conf.int = T)
```


## Alcohol-subgroups analysis - general population
```{r}
df6 <- df6 %>%
  mutate(
    alc_group = case_when(
      sex == 0 & alctot2 == 0 ~ 0,
      sex == 0 & alctot2 > 0 & alctot2 <= 10 ~ 1,
      sex == 0 & alctot2 > 10 ~ 2,
      sex == 1 & alctot2 == 0 ~ 0,
      sex == 1 & alctot2 > 0 & alctot2 <= 20 ~ 1,
      sex == 1 & alctot2 > 20 ~ 2,
      TRUE ~ NA_real_ 
    )
  )
table(df6$alc_group)

df_alc4grps <- df6 %>%
  mutate(
    alc_4grp = case_when(
      alctot2 == 0 ~ 0,
      alctot2 > 0 & alctot2*0.7 <= 8.4 ~ 1,
      alctot2*0.7 > 8.4 & alctot2*0.7 <= 15.4 ~ 2,
      alctot2*0.7 > 15.4 & alctot2*0.7 <= 24.5 ~ 3,
      alctot2*0.7 > 24.5 ~ 4,
      TRUE ~ NA_real_
    )
  )
table(df_alc4grps$alc_4grp)
```


```{r}
# LSBMD
lscat <- lm(std_lsbmd~as.factor(alc_group)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
tidy(lscat, conf.int = TRUE)

lscat_trend <- lm(std_lsbmd~as.numeric(alc_group)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(lscat_trend)
tidy(lscat_trend)
```
```{r}
# alcohol subgroups into 4 (0; 0-8.4d/wk; 8.4-15.4d/wk; 15.4-24.5d/wk; 24.5~d/wk)
ls4cat <- lm(std_lsbmd~as.factor(alc_4grp)+age+sex+income+education+smkstatus+PA_all+BMI,data = df_alc4grps)
tidy(ls4cat, conf.int = TRUE)

ls4cat_trend <- lm(std_lsbmd~as.numeric(alc_4grp)+age+sex+income+education+smkstatus+PA_all+BMI,data = df_alc4grps)
#summary(ls4cat_trend)
tidy(ls4cat_trend)
```



```{r}
# NKBMD
nkcat <- lm(std_fnbmd~as.factor(alc_group)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
tidy(nkcat, conf.int = TRUE)
nkcat_trend <- lm(std_fnbmd~as.numeric(alc_group)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(nkcat_trend)

tidy(nkcat_trend)
```
```{r}
# alcohol subgroups into 4 (0; 0-8.4d/wk; 8.4-15.4d/wk; 15.4-24.5d/wk; 24.5~d/wk)
nk4cat <- lm(std_fnbmd~as.factor(alc_4grp)+age+sex+income+education+smkstatus+PA_all+BMI,data = df_alc4grps)
tidy(nk4cat, conf.int = TRUE)

nk4cat_trend <- lm(std_fnbmd~as.numeric(alc_4grp)+age+sex+income+education+smkstatus+PA_all+BMI,data = df_alc4grps)
tidy(nk4cat_trend)
```



```{r}
# TOBMD
tocat <- lm(std_tobmd~as.factor(alc_group)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
tidy(tocat,conf.int = T)

tocat_trend <- lm(std_tobmd~as.numeric(alc_group)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(tocat_trend)
```
```{r}
# alcohol subgroups into 4 (0; 0-8.4d/wk; 8.4-15.4d/wk; 15.4-24.5d/wk; 24.5~d/wk)
to4cat <- lm(std_tobmd ~ as.factor(alc_4grp)+age+sex+income+education+smkstatus+PA_all+BMI,data = df_alc4grps)
tidy(to4cat, conf.int = TRUE)

to4cat_trend <- lm(std_tobmd~as.numeric(alc_4grp)+age+sex+income+education+smkstatus+PA_all+BMI,data = df_alc4grps)
tidy(to4cat_trend)
```


```{r}
# LSTBS
tbscat <- lm(std_lstbs~as.factor(alc_group)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
tidy(tbscat, conf.int=T)
tbscat_trend <- lm(std_lstbs~as.numeric(alc_group)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(tbscat_trend)

```
```{r}
# alcohol subgroups into 4 (0; 0-8.4d/wk; 8.4-15.4d/wk; 15.4-24.5d/wk; 24.5~d/wk)
tbs4cat <- lm(LSTBS~as.factor(alc_4grp)+age+sex+income+education+smkstatus+PA_all+BMI,data = df_alc4grps)
tidy(nk4cat, conf.int = TRUE)

nk4cat_trend <- lm(LSTBS~as.numeric(alc_4grp)+age+sex+income+education+smkstatus+PA_all+BMI,data = df_alc4grps)
tidy(nk4cat_trend)
```


## Forest plot - general population
```{r}
library(readxl)
fp <- read_excel("Forestplot_UKBB.xlsx")

# Forestplot
png("fp.png", width = 800, height = 600)
fp |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = .15, # We set the box size to better visualize the type
             line.margin = .1, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()
```

# Sex-stratified
```{r}
# setting datasets
df6_f <- df6[df6$sex == 0,]
df6_m <- df6[df6$sex == 1,]
```

## Female
```{r}
# LSBMD

lsmod1_f <- lm(std_lsbmd~log(alctot2/10+1)+age,data = df6_f)
lsmod2_f <- lm(std_lsbmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all,data = df6_f)
lsmod3_f <- lm(std_lsbmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)

tidy(lsmod1_f)
tidy(lsmod2_f)
tidy(lsmod3_f,conf.int=TRUE)
```

```{r}
#NKBMD

fnmod1_f <- lm(std_fnbmd~log(alctot2/10+1)+age,data = df6_f)
fnmod2_f <- lm(std_fnbmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all,data = df6_f)
fnmod3_f <- lm(std_fnbmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)

summary(fnmod1_f)
summary(fnmod2_f)
summary(fnmod3_f)

tidy(fnmod1_f)
tidy(fnmod2_f)
tidy(fnmod3_f,conf.int=TRUE)
```

```{r}
#TOBMD

tomod1_f <- lm(std_tobmd~log(alctot2/10+1)+age,data = df6_f)
tomod2_f <- lm(std_tobmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all,data = df6_f)
tomod3_f <- lm(std_tobmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)

summary(tomod1_f)
summary(tomod2_f)
tidy(tomod3_f,conf.int=TRUE)

```
```{r}
# LSTBS
tbsmod1_f <- lm(std_lstbs~log(alctot2/10+1)+age,data = df6_f)
tbsmod2_f <- lm(std_lstbs~log(alctot2/10+1)+age+income+education+smkstatus+PA_all,data = df6_f)
tbsmod3_f <- lm(std_lstbs~log(alctot2/10+1)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)

summary(tbsmod1_f)
summary(tbsmod2_f)
tidy(tbsmod3_f,conf.int=TRUE)
```

## Alcohol-subgroups analysis - Female
```{r}
table(df6_f$alc_group)
```

```{r}
# LSBMD
lscat_f <- lm(std_lsbmd~as.factor(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
tidy(lscat_f, conf.int = T)

lscat_f_trend <- lm(std_lsbmd~as.numeric(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(lscat_f_trend)
tidy(lscat_f_trend)
```

```{r}
# NKBMD
nkcat_f <- lm(std_fnbmd~as.factor(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
tidy(nkcat_f, conf.int = T)

nkcat_f_trend <- lm(std_fnbmd~as.numeric(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(nkcat_f_trend)


tidy(nkcat_f_trend)
```


```{r}
# TOBMD
tocat_f <- lm(std_tobmd~as.factor(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
tidy(tocat_f, conf.int = T)

tocat_f_trend <- lm(std_tobmd~as.numeric(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(tocat_f_trend)
```
```{r}
# LSTBS
tbscat_f <- lm(std_lstbs~as.factor(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
tidy(tbscat_f, conf.int = T)

```

## Forest plot - female
```{r}

fp_f <- read_excel("Forestplot_F.xlsx")

# Forestplot - Female
png("fp_f.png", width = 800, height = 600)
fp_f |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = .15, # We set the box size to better visualize the type
             line.margin = .1, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()
```

## Male
```{r}
# LSBMD
lsmod1_m <- lm(std_lsbmd~log(alctot2/10+1)+age,data = df6_m)
lsmod2_m <- lm(std_lsbmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all,data = df6_m)
lsmod3_m <- lm(std_lsbmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)

summary(lsmod1_m)
summary(lsmod2_m)
summary(lsmod3_m)

tidy(lsmod1_m)
tidy(lsmod2_m)
tidy(lsmod3_m,conf.int = TRUE)

```

```{r}
#NKBMD

fnmod1_m <- lm(std_fnbmd~log(alctot2/10+1)+age,data = df6_m)
fnmod2_m <- lm(std_fnbmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all,data = df6_m)
fnmod3_m <- lm(std_fnbmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)

summary(fnmod1_m)
summary(fnmod2_m)
summary(fnmod3_m)

tidy(fnmod1_m)
tidy(fnmod2_m)
tidy(fnmod3_m,conf.int = TRUE)

```

```{r}
#TOBMD
tomod1_m <- lm(std_tobmd~log(alctot2/10+1)+age,data = df6_m)
tomod2_m <- lm(std_tobmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all,data = df6_m)
tomod3_m <- lm(std_tobmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)

tidy(tomod1_m)
tidy(tomod2_m)
tidy(tomod3_m,conf.int = TRUE)
```
```{r}
# LSTBS
tbsmod1_m <- lm(std_tobmd~log(alctot2/10+1)+age,data = df6_m)
tbsmod2_m <- lm(std_tobmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all,data = df6_m)
tbsmod3_m <- lm(std_tobmd~log(alctot2/10+1)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)

tidy(tbsmod1_m)
tidy(tbsmod2_m)
tidy(tbsmod3_m,conf.int = TRUE)
```

## Alcohol-subgroups analysis - Male
```{r}
table(df6_m$alc_group)
```

```{r}
# LSBMD
lscat_m <- lm(std_lsbmd~as.factor(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
tidy(lscat_m,conf.int = T)

lscat_m_trend <- lm(std_lsbmd~as.numeric(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(lscat_m_trend)


tidy(lscat_m_trend)
```

```{r}
# NKBMD
nkcat_m <- lm(std_fnbmd~as.factor(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
tidy(nkcat_m, conf.int = T)

nkcat_m_trend <- lm(std_fnbmd~as.numeric(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(nkcat_m_trend)


tidy(nkcat_m_trend)
```

```{r}
# TOBMD
tocat_m <- lm(std_tobmd~as.factor(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
tidy(tocat_m,conf.int = T)

tocat_m_trend <- lm(std_tobmd~as.numeric(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(tocat_m_trend)
```

```{r}
#LSTBS
tbscat_m <- lm(std_lstbs~as.factor(alc_group)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
tidy(tbscat_m,conf.int = T)
```

# Forest plot - male
```{r}
fp_m <- read_excel("Forestplot_M.xlsx")
# Forestplot - Female
png("fp_m.png", width = 800, height = 600)
fp_m |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = .15, # We set the box size to better visualize the type
             line.margin = .1, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()

```

# Age-stratified analysis
```{r}
df7 <- df7 %>% 
  group_by(sex) %>% 
  mutate(std_lstbs = as.numeric(scale(LSTBS))) %>% 
  ungroup()

df7_old <- subset(df7, age>= 60)
df7_young <- subset(df7, age < 60)
```

## Age < 60yrs
```{r}
# lsbmd
lsmod3_young <- lm(std_lsbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df7_young)
tidy(lsmod3_young,conf.int=TRUE)
```
```{r}
# fnbmd
fnmod3_young <- lm(std_fnbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df7_young)
tidy(fnmod3_young,conf.int=TRUE)
```
```{r}
# tobmd
tomod3_young <- lm(std_tobmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df7_young)
tidy(tomod3_young,conf.int=TRUE)
```
```{r}
#lstbs
tbsmod3_young <- lm(std_lstbs~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df7_young)
tidy(tbsmod3_young,conf.int=TRUE)
```
## Age >= 60yrs
```{r}
# lsbmd
lsmod3_old <- lm(std_lsbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df7_old)
tidy(lsmod3_old,conf.int=TRUE)
```
```{r}
# fnbmd
fnmod3_old <- lm(std_fnbmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df7_old)
tidy(fnmod3_old,conf.int=TRUE)
```
```{r}
# tobmd
tomod3_old <- lm(std_tobmd~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df7_old)
tidy(tomod3_old,conf.int=TRUE)
```
```{r}
#lstbs
tbsmod3_old <- lm(std_lstbs~log(alctot2/10+1)+age+sex+income+education+smkstatus+PA_all+BMI,data = df7_old)
tidy(tbsmod3_old,conf.int=TRUE)
```

# Results for 5 quintiles of alcohol
```{r}
# 5 quintiles
deciles <- quantile(df6$alctot2, probs = seq(0, 1, by=0.2))
deciles
quintile_f <- quantile(df6_f$alctot2, probs = seq(0,1,by=0.2))
quintile_m <- quantile(df6_m$alctot2, probs = seq(0,1,by=0.2))

df6 <- df6 %>%
  mutate(alccat = cut(alctot2,breaks = deciles,include.lowest = T))
df6_f <- df6_f %>%
  mutate(alccat_f = cut(alctot2, breaks = quintile_f, include.lowest = T))
df6_m <- df6_m %>%
  mutate(alccat_m = cut(alctot2, breaks = quintile_m, include.lowest = T))
```

## General population
```{r}
table(df6$alccat)
```

```{r}
lscat4 <- lm(std_lsbmd~alccat+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(lscat4)
confint(lscat4)
lscat4 <- lm(std_lsbmd~as.numeric(alccat)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(lscat4)
```

```{r}
fncat4 <- lm(std_fnbmd~alccat+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(fncat4)
confint(fncat4)
fncat4 <- lm(std_fnbmd~as.numeric(alccat)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(fncat4)
```

```{r}
tocat4 <- lm(std_tobmd~alccat+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(tocat4)
confint(tocat4)
tocat4 <- lm(std_tobmd~as.numeric(alccat)+age+sex+income+education+smkstatus+PA_all+BMI,data = df6)
summary(tocat4)
```

## Forest plot of 5 quintiles - general population
```{r}
fp_5q <- read_excel("Forestplot_5q.xlsx")
# Forestplot
fp_5q1 <- ggplot(fp_5q,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.2)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=2.0)
fp_5q1
# ggsave("fp_5q.png",width = 6,height = 5)
```

## Female
```{r}
table(df6_f$alccat_f)
```

```{r}
lscat4_f <- lm(std_lsbmd~alccat_f+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(lscat4_f)
confint(lscat4_f)
lscat4_f <- lm(std_lsbmd~as.numeric(alccat_f)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(lscat4_f)
```


```{r}
fncat4_f <- lm(std_fnbmd~alccat_f+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(fncat4_f)
confint(fncat4_f)
fncat4_f <- lm(std_fnbmd~as.numeric(alccat_f)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(fncat4)
```

```{r}
tocat4_f <- lm(std_tobmd~alccat_f+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(tocat4_f)
confint(tocat4_f)
tocat4_f <- lm(std_tobmd~as.numeric(alccat_f)+age+income+education+smkstatus+PA_all+BMI,data = df6_f)
summary(tocat4_f)
```

## Forest plot of 5 q - female
```{r}
fp_5q_f <- read_excel("Forestplot_5q_F.xlsx")
# Forestplot
fp_5q_f1 <- ggplot(fp_5q_f,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.2)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=2.0)
fp_5q_f1
#ggsave("fp_5q_f.png",width = 6,height = 5)
```

## Male
```{r}
table(df6_m$alccat_m)
```

```{r}
lscat4_m <- lm(std_lsbmd~alccat_m+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(lscat4_m)
confint(lscat4_m)
lscat4_m <- lm(std_lsbmd~as.numeric(alccat_m)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(lscat4_m)
```

```{r}
fncat4_m <- lm(std_fnbmd~alccat_m+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(fncat4_m)
confint(fncat4_m)
fncat4_m <- lm(std_fnbmd~as.numeric(alccat_m)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(fncat4)
```

```{r}
tocat4_m <- lm(std_tobmd~alccat_m+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(tocat4_m)
confint(tocat4_m)
tocat4_m <- lm(std_tobmd~as.numeric(alccat_m)+age+income+education+smkstatus+PA_all+BMI,data = df6_m)
summary(tocat4_m)
```
## Forest plot of 5 q - male
```{r}
# Forestplot
fp_5q_m <- read_excel("Forestplot_5q_M.xlsx")
# Forestplot
fp_5q_m1 <- ggplot(fp_5q_m,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.2)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=2.0)
fp_5q_m1
#ggsave("fp_5q_m.png",width = 6,height = 5)
```

# One-sample MR
```{r}
# 78 SNPs involved in the analysis in UKBB

prs_alc <- read.table("ALC_PRS_99.all_score",header = T) 
prs_alc <- prs_alc %>%
  rename(eid = IID) 
prs_alc <- prs_alc[,-1]

PRS_ALC_sum_78 <- read.table("//store/department/im/share/personal/059909/PHD/Bone & Alcohol/SNPs_RS_and_UKB/PRS_UKB/PRS_78SNPs/ALC_PRS_sum_matchinbothcohorts.all_score", header = TRUE)
PRS_ALC_sum_78 <- PRS_ALC_sum_78 %>%
  select(IID, Pt_5e.08) %>%
  rename(eid = IID, PRS_sum_78 = Pt_5e.08)


PC_ukb <- read.csv("PC_ukb.csv", header = T)

PRS_ALC_96_useProxy <- read.table("ALC_PRS_sum_Allcohort_keepAmbig_useProxy.all_score", header = TRUE)
PRS_ALC_96_useProxy <- PRS_ALC_96_useProxy %>%
  select(IID, Pt_5e.08) %>%
  rename(eid=IID, PRS_sum_96=Pt_5e.08)

df7 <- left_join(df6, prs_alc, by = "eid")
df7 <- left_join(df7, PC_ukb, by = "eid")
df7 <- left_join(df7, PRS_ALC_sum_78, by = "eid")
df7 <- left_join(df7, ukb_lstbs_origin, by = "eid")

df7 <- df7 %>% 
  group_by(sex) %>% 
  mutate(std_lsbmd = as.vector(scale(lsbmd)),
         std_fnbmd = as.vector(scale(fnbmd)),
         std_tobmd = as.vector(scale(tobmd))) %>% 
  ungroup()

df7 <- left_join(df7,PRS_ALC_96_useProxy, by = "eid")

PCs <- paste("PC",1:10,sep = "")
all_vars <- c("log(alctot2/10+1)","age","sex",PCs)
all_ivs <- c("PRS_sum_78","age","sex",PCs)
all_ivs_77 <- c("PGS_77","age","sex",PCs)
all_ivs_withoutUKB_77 <- c("PGS_withoutUKB_77","age","sex",PCs)

# Histogram of PRS of alcohol consumption
hist(df7$Pt_5e.08, main = "Histogram of PRS of alcohol consumption", xlab = "PRS of alcohol consumption")
```

## Input datasets 77
```{r}
prs_alc_77 <- read.table("PRS_ALC_sum_77.all_score",header = T) 
prs_alc_77 <- prs_alc_77 %>%
  rename(eid = IID) %>% 
  rename(PGS_77 = Pt_5e.08)
prs_alc_77 <- prs_alc_77[,-1]
```

## Input datasets 77 withoutUKB
```{r}
prs_alc_withoutUKB_77 <- read.table("ALC_PRS_sum_Allcohort_withoutUKB_77.all_score",header = T) 
prs_alc_withoutUKB_77 <- prs_alc_withoutUKB_77 %>%
  rename(eid = IID) %>% 
  rename(PGS_withoutUKB_77 = Pt_5e.08) %>% 
  select(eid, PGS_withoutUKB_77)
```


```{r}  
df7 <- left_join(df7, prs_alc_77, by="eid")
df7 <- left_join(df7, prs_alc_withoutUKB_77, by="eid")
```

```{r}
# LSBMD
formula_string_ls <- paste("std_lsbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs, collapse = "+"))
formula_ls <- as.formula(formula_string_ls)
formula_ls
mrmod_ls <- ivreg(formula_ls, data = df7)
tidy(mrmod_ls)
```

```{r}
mrmod_ls_proxy <- ivreg(std_lsbmd ~ log(alctot2/10 + 1) + age + sex + PC1 + PC2 + PC3 + 
    PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 | PRS_sum_96 + age + 
    sex + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + 
    PC10, data = df7)
tidy(mrmod_ls_proxy,conf.int = TRUE)
```

```{r}
# 77
formula_string_ls_77 <- paste("std_lsbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_77, collapse = "+"))
formula_string_ls_77 <- as.formula(formula_string_ls_77)
formula_string_ls_77
mrmod_ls_77 <- ivreg(formula_string_ls_77, data = df7)
summary(mrmod_ls_77)
```
```{r}
# 77 without UKB
formula_string_ls_withoutUKB_77 <- paste("std_lsbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_withoutUKB_77, collapse = "+"))
formula_string_ls_withoutUKB_77 <- as.formula(formula_string_ls_withoutUKB_77)
formula_string_ls_withoutUKB_77
mrmod_ls_withoutUKB_77 <- ivreg(formula_string_ls_withoutUKB_77, data = df7)
summary(mrmod_ls_withoutUKB_77)
```

```{r}
# FNBMD
formula_string_fn <- paste("std_fnbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs, collapse = "+"))
formula_fn <- as.formula(formula_string_fn)
formula_fn
mrmod_fn <- ivreg(formula_fn, data = df7)
tidy(mrmod_fn)
```
```{r}
mrmod_fn_proxy <- ivreg(std_fnbmd ~ log(alctot2/10 + 1) + age + sex + PC1 + PC2 + PC3 + 
    PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 | PRS_sum_96 + age + 
    sex + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + 
    PC10, data = df7)
tidy(mrmod_fn_proxy, conf.int = TRUE)
```

```{r}
# 77
formula_string_fn_77 <- paste("std_fnbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_77, collapse = "+"))
formula_string_fn_77 <- as.formula(formula_string_fn_77)
formula_string_fn_77
mrmod_fn_77 <- ivreg(formula_string_fn_77, data = df7)
summary(mrmod_fn_77)
```
```{r}
# 77 without UKB
formula_string_fn_withoutUKB_77 <- paste("std_fnbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_withoutUKB_77, collapse = "+"))
formula_string_fn_withoutUKB_77 <- as.formula(formula_string_fn_withoutUKB_77)
formula_string_fn_withoutUKB_77
mrmod_fn_withoutUKB_77 <- ivreg(formula_string_fn_withoutUKB_77, data = df7)
summary(mrmod_fn_withoutUKB_77)
```

```{r}
# TOBMD
formula_string_to <- paste("std_tobmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs, collapse = "+"))
formula_to <- as.formula(formula_string_to)
formula_to
mrmod_to <- ivreg(formula_to, data = df7)
tidy(mrmod_to)
```
```{r}
mrmod_to_proxy <- ivreg(std_tobmd ~ log(alctot2/10 + 1) + age + sex + PC1 + PC2 + PC3 + 
    PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 | PRS_sum_96 + age + 
    sex + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + 
    PC10, data = df7)
tidy(mrmod_to_proxy, conf.int=TRUE)
```

```{r}
# 77
formula_string_to_77 <- paste("std_tobmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_77, collapse = "+"))
formula_string_to_77 <- as.formula(formula_string_to_77)
formula_string_to_77
mrmod_to_77 <- ivreg(formula_string_to_77, data = df7)
summary(mrmod_to_77)
```
```{r}
# 77 without UKB
formula_string_to_withoutUKB_77 <- paste("std_tobmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_withoutUKB_77, collapse = "+"))
formula_string_to_withoutUKB_77 <- as.formula(formula_string_to_withoutUKB_77)
formula_string_to_withoutUKB_77
mrmod_to_withoutUKB_77 <- ivreg(formula_string_to_withoutUKB_77, data = df7)
summary(mrmod_to_withoutUKB_77)
```

```{r}
# LSTBS
formula_string_tbs <- paste("LSTBS ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs, collapse = "+"))
formula_tbs <- as.formula(formula_string_tbs)
formula_tbs
mrmod_tbs <- ivreg(formula_tbs, data = df7)
tidy(mrmod_tbs)
```
```{r}
mrmod_tbs_proxy <- ivreg(LSTBS ~ log(alctot2/10 + 1) + age + sex + PC1 + PC2 + PC3 + PC4 + 
    PC5 + PC6 + PC7 + PC8 + PC9 + PC10 | PRS_sum_96 + age + sex + 
    PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10, data = df7)
tidy(mrmod_tbs_proxy, conf.int = TRUE)
```

```{r}
# 77
formula_string_tbs_77 <- paste("LSTBS ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_77, collapse = "+"))
formula_string_tbs_77 <- as.formula(formula_string_tbs_77)
formula_string_tbs_77
mrmod_tbs_77 <- ivreg(formula_string_tbs_77, data = df7)
summary(mrmod_tbs_77)
```
```{r}
# 77 without UKB
formula_string_tbs_withoutUKB_77 <- paste("LSTBS ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_withoutUKB_77, collapse = "+"))
formula_string_tbs_withoutUKB_77 <- as.formula(formula_string_tbs_withoutUKB_77)
formula_string_tbs_withoutUKB_77
mrmod_tbs_withoutUKB_77 <- ivreg(formula_string_tbs_withoutUKB_77, data = df7)
summary(mrmod_tbs_withoutUKB_77)
```

# Sex-stratified
```{r}
PCs1 <- paste("PC",1:20,sep = "")
all_vars_withouteth1 <-  c("log(alctot2/10+1)","age",PCs1)
all_ivs_withouteth1 <- c("Pt_5e.08","age","sex",PCs1)
all_ivs_withouteth2 <- c("Pt_5e.08","age",PCs1)
```
## Female
```{r}
df7_f <- subset(df7,df7$sex == 0)
df7_m <- subset(df7,df7$sex == 1)

formula_string_ls3 <- paste("std_lsbmd ~", paste(all_vars_withouteth1, collapse = "+"),"|",paste(all_ivs_withouteth2, collapse = "+"))
formula_string_fn3 <- paste("std_fnbmd ~", paste(all_vars_withouteth1, collapse = "+"),"|",paste(all_ivs_withouteth2, collapse = "+"))
formula_string_to3 <- paste("std_tobmd ~", paste(all_vars_withouteth1, collapse = "+"),"|",paste(all_ivs_withouteth2, collapse = "+"))
formula_string_tbs3 <- paste("LSTBS ~ ", paste(all_vars_withouteth1, collapse = "+"),"|",paste(all_ivs_withouteth2, collapse = "+"))

formula_ls3 <- as.formula(formula_string_ls3)
formula_fn3 <- as.formula(formula_string_fn3)
formula_to3 <- as.formula(formula_string_to3)
formula_tbs3 <- as.formula(formula_string_tbs3)

formula_string_ls4 <- paste("std_lsbmd ~ sex +", paste(all_vars_withouteth1, collapse = "+"),"|",paste(all_ivs_withouteth1, collapse = "+"))
formula_string_fn4 <- paste("std_fnbmd ~ sex +", paste(all_vars_withouteth1, collapse = "+"),"|",paste(all_ivs_withouteth1, collapse = "+"))
formula_string_to4 <- paste("std_tobmd ~ sex +", paste(all_vars_withouteth1, collapse = "+"),"|",paste(all_ivs_withouteth1, collapse = "+"))
formula_string_tbs4 <- paste("LSTBS ~ sex +", paste(all_vars_withouteth1, collapse = "+"),"|",paste(all_ivs_withouteth1, collapse = "+"))

formula_ls4 <- as.formula(formula_string_ls4)
formula_fn4 <- as.formula(formula_string_fn4)
formula_to4 <- as.formula(formula_string_to4)
formula_tbs4 <- as.formula(formula_string_tbs4)
```

```{r}
mr_ls_f <- ivreg(formula_ls3, data = df7_f)
tidy(mr_ls_f,conf.int=T)
```

```{r}
mr_fn_f <- ivreg(formula_fn3, data = df7_f)
tidy(mr_fn_f,conf.int = T)
```


```{r}
mr_to_f <- ivreg(formula_to3, data = df7_f)
tidy(mr_to_f, conf.int = T)
```
```{r}
mr_tbs_f <- ivreg(formula_tbs3, data = df7_f)
tidy(mr_tbs_f, conf.int = TRUE)
```

## Male
```{r}
mr_ls_m <- ivreg(formula_ls3, data = df7_m)
tidy(mr_ls_m, conf.int = TRUE)
```
```{r}
mr_fn_m <- ivreg(formula_fn3, data = df7_m)
tidy(mr_fn_m, conf.int = TRUE)
```
```{r}
mr_to_m <- ivreg(formula_to3, data = df7_m)
tidy(mr_to_m, conf.int = TRUE)
```
```{r}
mr_tbs_m <- ivreg(formula_tbs3, data = df7_m)
tidy(mr_tbs_m, conf.int = TRUE)
```
## Female over 55 year-old
```{r}
df9_f <- subset(df7_f, age >= 55)

mr_ls_f_over55 <- ivreg(formula_ls3, data = df9_f)
summary(mr_ls_f_over55)

mr_fn_f_over55 <- ivreg(formula_fn3, data = df9_f)
summary(mr_fn_f_over55)

mr_to_f_over55 <- ivreg(formula_to3, data = df9_f)
summary(mr_to_f_over55)

```

## Age-stratified MR
```{r}
df7_young <- subset(df7, age < 60)
df7_old <- subset(df7, age >= 60)
```
## Young
```{r}
mr_ls_young <- ivreg(formula_ls4, data = df7_young)
tidy(mr_ls_young, conf.int = TRUE)

mr_fn_young <- ivreg(formula_fn4, data = df7_young)
tidy(mr_fn_young, conf.int=TRUE)

mr_to_young <- ivreg(formula_to4, data = df7_young)
tidy(mr_to_young,conf.int = TRUE)

mr_tbs_young <- ivreg(formula_tbs4, data = df7_young)
tidy(mr_tbs_young, conf.int = TRUE)
```
## Old
```{r}
mr_ls_old <- ivreg(formula_ls4, data = df7_old)
tidy(mr_ls_old, conf.int = TRUE)

mr_fn_old <- ivreg(formula_fn4, data = df7_old)
tidy(mr_fn_old, conf.int=TRUE)

mr_to_old <- ivreg(formula_to4, data = df7_old)
tidy(mr_to_old,conf.int = TRUE)

mr_tbs_old <- ivreg(formula_tbs4, data = df7_old)
tidy(mr_tbs_old, conf.int = TRUE)

```


# filter out outliers
```{r}
# filter out lsbmd > 4 st.d.
df7$lsbmd_zscore <- (df7$lsbmd - mean(df6$lsbmd)) / sd(df7$lsbmd)

df7$ls_outlier <- if_else(abs(df7$lsbmd_zscore) > 4, 1 ,0)
table(df7$ls_outlier)
```

```{r}
df8 <- subset(df7, ls_outlier == 0)
summary(df8$lsbmd)
```

```{r}
formula_string_ls <- paste("std_lsbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs, collapse = "+"))
formula_ls_withoutOutliers <- as.formula(formula_string_ls)
formula_ls_withoutOutliers
mrmod_ls_withoutOutliers <- ivreg(formula_ls_withoutOutliers, data = df8)
summary(mrmod_ls_withoutOutliers)
```

# Input LS-TBS data
```{r}
ukb_lstbs_origin <- read.csv("//store/department/im/share/projects/Fernando_Group/Alcohol_BMD_Xinpeng/UK Biobank/datasets/ukbb_TBS_participant (KT extraction).csv", header = TRUE)

# input height and weight for QC
ukb_weight_and_height <- read.csv("//store/department/im/share/projects/Fernando_Group/Alcohol_BMD_Xinpeng/UK Biobank/datasets/weight_and_height.csv", header = TRUE)

names(ukb_lstbs_origin)[2] <- "LSTBS"
ukb_weight_and_height <- ukb_weight_and_height %>%
  select(eid, X12143.2.0, X12144.2.0) %>%
  rename(weight = X12143.2.0, height = X12144.2.0)

# Combine LSTBS, weight, and height
ukb_lstbs_weight_height <- left_join(ukb_lstbs_origin, ukb_weight_and_height, by = "eid")
```

# Histrogram and Correlation plots
```{r}
# Histogram
hist_ukb_lstbs <- ggplot(ukb_lstbs_weight_height, aes(x = LSTBS)) + 
  geom_histogram(binwidth = 0.05, fill = "grey", color = "black", alpha = 0.7) +
  labs(title = "Histogram of LSTBS in UKBB", x = "LSTBS", y = "Frequency") +
  theme_minimal()
hist_ukb_lstbs
ggsave("hist_ukb_lstbs.png", width = 6, height = 5)
```

```{r}
# Correlation plot - Weight and LSTBS
corplot_ukb_lstbs_weight <- ggplot(ukb_lstbs_weight_height, aes(x = LSTBS, y = weight)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", color = "red") +
  stat_cor(method = "pearson") +
  labs(title = "Correlation between Weight and LSTBS",
       x = "Weight(kg)",
       y = "LSTBS") +
  theme(panel.grid = element_blank())
corplot_ukb_lstbs_weight
ggsave("Correlation plot - Weight_and_LSTBS.png", width = 6, height = 5)
```

```{r}
# Correlation plot - Height and LSTBS
corplot_ukb_lstbs_height <- ggplot(ukb_lstbs_weight_height, aes(x = LSTBS, y = height)) +
  geom_point(color = "black", size = 3) +
  geom_smooth(method = "lm", color = "red") +
  stat_cor(method = "pearson") + 
  labs(title = "Correlation between Height and LSTBS",
       x = "Height(cm)",
       y = "LSTBS") +
  theme(panel.grid = element_blank())
corplot_ukb_lstbs_height
ggsave("Correlation plot - Height_and_LSTBS.png", width = 6, height = 5)
```

# Input PRS for PAU in UKBB
```{r}
PAU_PRS_sum <- read.table("PRS_PAU_sum.all_score", header = T)
PAU_PRS_sum <- PAU_PRS_sum %>% 
  rename(eid = IID,
         wPRS_PAU = Pt_5e.08) %>% 
  select(eid, wPRS_PAU)

df_PAU <- left_join(df7, PAU_PRS_sum, by="eid")

all_ivs_PAU <- c("wPRS_PAU",all_ivs[-1])

PAU_PRS_sum_18 <- read.table("PRS_PAU_sum_18.all_score", header = T)

PAU_PRS_sum_18 <- PAU_PRS_sum_18 %>% 
  rename(eid = IID,
         wPRS_PAU_18 = Pt_5e.08) %>% 
  select(eid, wPRS_PAU_18)

df_PAU <- left_join(df_PAU, PAU_PRS_sum_18, by="eid")
all_ivs_PAU_18 <- c("wPRS_PAU_18",all_ivs_PAU[-1])

```


```{r}
# LSBMD
formula_string_ls_PAU <- paste("std_lsbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_PAU_18, collapse = "+"))
formula_ls_PAU <- as.formula(formula_string_ls_PAU)
formula_ls_PAU
mrmod_ls_PAU <- ivreg(formula_ls_PAU, data = df_PAU)
tidy(mrmod_ls_PAU, conf.int = T)
```

```{r}
# FNBMD
formula_string_fn_PAU <- paste("std_fnbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_PAU_18, collapse = "+"))
formula_fn_PAU <- as.formula(formula_string_fn_PAU)
formula_fn_PAU
mrmod_fn_PAU <- ivreg(formula_fn_PAU, data = df_PAU)
tidy(mrmod_fn_PAU, conf.int = T)
```

```{r}
# TOBMD
formula_string_to_PAU <- paste("std_tobmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_PAU_18, collapse = "+"))
formula_to_PAU <- as.formula(formula_string_to_PAU)
formula_to_PAU
mrmod_to_PAU <- ivreg(formula_to_PAU, data = df_PAU)
tidy(mrmod_to_PAU,conf.int=T)
```
```{r}
# LSTBS
formula_string_tbs_PAU <- paste("LSTBS ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_PAU_18, collapse = "+"))
formula_tbs_PAU <- as.formula(formula_string_tbs_PAU)
formula_tbs_PAU
mrmod_tbs_PAU <- ivreg(formula_tbs_PAU, data = df_PAU)
tidy(mrmod_tbs_PAU,conf.int=T)
```


# Addtional analysis using SNPs without UKB
```{r}
PRS_withoutUKB <- read.table("ALC_PRS_sum_Allcohort_withoutUKB.all_score", header = T)
PRS_withoutUKB <- PRS_withoutUKB %>% 
  rename(eid = IID,
         wPRS_withoutUKB = Pt_5e.08) %>% 
  select(eid, wPRS_withoutUKB)

df_PRS_withoutUKB <- left_join(df7, PRS_withoutUKB, by="eid")

all_ivs_withoutUKB <- c("wPRS_withoutUKB",all_ivs[-1])

```

# One-sample MR
```{r}
# LSBMD
formula_string_ls_withoutUKB <- paste("std_lsbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_withoutUKB, collapse = "+"))
formula_ls_withoutUKB <- as.formula(formula_string_ls_withoutUKB)
formula_ls_withoutUKB
mrmod_ls_withoutUKB <- ivreg(formula_ls_withoutUKB, data = df_PRS_withoutUKB)
tidy(mrmod_ls_withoutUKB, conf.int = T)
```

```{r}
# FNBMD
formula_string_fn_withoutUKB <- paste("std_fnbmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_withoutUKB, collapse = "+"))
formula_fn_withoutUKB <- as.formula(formula_string_fn_withoutUKB)
formula_fn_withoutUKB
mrmod_fn_withoutUKB <- ivreg(formula_fn_withoutUKB, data = df_PRS_withoutUKB)
tidy(mrmod_fn_withoutUKB, conf.int = T)
```

```{r}
# TOBMD
formula_string_to_withoutUKB <- paste("std_tobmd ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_withoutUKB, collapse = "+"))
formula_to_withoutUKB <- as.formula(formula_string_to_withoutUKB)
formula_to_withoutUKB
mrmod_to_withoutUKB <- ivreg(formula_to_withoutUKB, data = df_PRS_withoutUKB)
tidy(mrmod_to_withoutUKB,conf.int=T)
```

```{r}
# LSTBS
formula_string_tbs_withoutUKB <- paste("LSTBS ~", paste(all_vars, collapse = "+"),"|",paste(all_ivs_withoutUKB, collapse = "+"))
formula_tbs_withoutUKB <- as.formula(formula_string_tbs_withoutUKB)
formula_tbs_withoutUKB
mrmod_tbs_withoutUKB <- ivreg(formula_tbs_withoutUKB, data = df_PRS_withoutUKB)
tidy(mrmod_tbs_withoutUKB,conf.int=T)
```