---
title: "ALC_BONE_RS"
author: "Xinpeng"
date: "2026-03-22"
output: html_document
---

# Settings
```{r}
#setwd("//store/department/im/share/personal/059909/PHD/Bone & Alcohol/ERGO5/RS_E5")

# Library packages
library(tidyverse)   # covers dplyr, ggplot2, tidyr
library(haven)
library(eeptools)
library(sjlabelled)
library(readxl)
library(AER)
library(tableone)
library(Hmisc)
library(broom)
library(forestplot)
```

# Input original datasets
```{r}
RoterdamStudy_Basics2014 <- read_sav("RoterdamStudy_Basics2014.sav")
RS_GeneticEtnicity_11_JAN_2016_ <- read_sav("RS_GeneticEtnicity_(11-JAN-2016).sav")
e5_intvw_Alcoholperday_05_11_2019 <- read_sav("e5_intvw_Alcoholperday_05-11-2019.sav")
e5_intvw_SMOKING_22_mar_2021_ <- read_sav("e5_intvw_SMOKING_(22-mar-2021).sav")
Ergo5_RSI_5_RSII_3_RSIII_2_Femur_DXA_iDXA_BMD <- read_sav("Ergo5_(RSI-5_RSII-3_RSIII-2)_Femur_DXA_iDXA_BMD.sav")
e5_5_ls_bmd_ls_tbs_15_feb_2022_ <- read_sav("e5_(5)_ls.bmd ls.tbs_(15-feb-2022).sav")
RS_GeneticEtnicity_11_JAN_2016_ <- read_sav("RS_GeneticEtnicity_(11-JAN-2016).sav")
e5_intvw_SES_07_feb_2014_ <- read_sav("e5_intvw_SES_(07-feb-2014).sav")
e_InformedConsent_24_MAY_2018_ <- read_sav("e_InformedConsent_(24-MAY-2018).sav")
Ergo5_RSI_5_RSII_3_RSIII_2_Total_Body_DXA_iDXA_BMD <- read_sav("Ergo5_(RSI-5_RSII-3_RSIII-2)_Total_Body_DXA_iDXA_BMD.sav")
e1epej_SES <- read_sav("e1epej_SES.sav")
Education_RS_I_II_III_UNESCO_class_12_MAR_2015_ <- read_sav("Education RS-I-II-III (UNESCO class)_(12-MAR-2015).sav")
e5_5_ANTHROPO_10_dec_2015_ <- read_sav("e5_(5)_ANTHROPO_(10-dec-2015).sav")
ERGO_5_RS_I_5_II_3_III_2_LASA_07_2020_ <- read_sav("ERGO 5 (RS-I-5, II-3, III-2) - LASA (07-2020).sav")
relateness_0125 <- read.table("//store/department/im/share/projects/Fernando_Group/Alcohol_BMD_Xinpeng/Rotterdam Study/related_0125.txt", header = T)
```

# Create datasets for analysis
```{r}
Basic_ergo5 <- RoterdamStudy_Basics2014
Basic_ergo5 <- left_join(Basic_ergo5,RS_GeneticEtnicity_11_JAN_2016_,by="ergoid")

e5_InformConsent <- e_InformedConsent_24_MAY_2018_[,c(1,32)]
e5_InformConsent <- e5_InformConsent %>%
  rename(e5_informconsent = e5_4109) %>%
  mutate(e5_informconsent = na_if(e5_informconsent, 8)) %>%
  mutate(e5_informconsent = na_if(e5_informconsent, 9)) %>%
  mutate(e5_informconsent = case_when(
    e5_informconsent == 1 ~ 1,
    e5_informconsent == 2 ~ 1,
    e5_informconsent == 4 ~ 0,
    TRUE ~ e5_informconsent
  ))

dataset1 <- left_join(Basic_ergo5,e5_InformConsent, by ="ergoid")
dataset1 <- dataset1[!is.na(dataset1$e5_informconsent),]
dataset1 <- dataset1[dataset1$e5_informconsent == 1,] # in total 7146 participants have inform consent

# outcomes & exposures & covariates
e5_alc <- e5_intvw_Alcoholperday_05_11_2019[,c(1,21:24)]
e5_ls_bmd_tbs <- e5_5_ls_bmd_ls_tbs_15_feb_2022_[,c(1,6,13,14)]
e5_fnbmd <- Ergo5_RSI_5_RSII_3_RSIII_2_Femur_DXA_iDXA_BMD[,c(1,3,6,9)]
e5_tobmd <- Ergo5_RSI_5_RSII_3_RSIII_2_Total_Body_DXA_iDXA_BMD[,c(1,6)]
e5_SES <- e1epej_SES[,c(1,6)]
e5_smk <- e5_intvw_SMOKING_22_mar_2021_[,c(1,7)]
e5_education <- Education_RS_I_II_III_UNESCO_class_12_MAR_2015_[,c(1,3)]
e5_PA <- ERGO_5_RS_I_5_II_3_III_2_LASA_07_2020_[,c(1,3,4)]
e5_BMI <- e5_5_ANTHROPO_10_dec_2015_[,c(1,6,7)] %>%
  mutate(BMI = e5_230/(e5_229/100)^2)

e5_smk <- e5_intvw_SMOKING_22_mar_2021_ %>%
  mutate(smk_status = case_when(
    e5_EILF6 == 0 & e5_EILFE==0 & e5_EILF4==0 & e5_EILF5==0 ~ 0,
    (e5_EILF4==1|e5_EILF4==2|e5_EILF4==3)|e5_EILF5==1|(e5_EILF6==0&e5_EILFE==1) ~ 1,
    (e5_EILF4==4|e5_EILF4==5|e5_EILF4==6)|e5_EILF5==2|e5_EILF5==3|e5_EILF6==1 ~ 2
  )) %>%
  select(ergoid, smk_status)
  

dataset2 <- left_join(dataset1, e5_BMI, by = "ergoid")
dataset2 <- left_join(dataset2, e5_alc, by = "ergoid")
dataset2 <- left_join(dataset2, e5_ls_bmd_tbs, by = "ergoid")
dataset2 <- left_join(dataset2, e5_fnbmd, by = "ergoid")
dataset2 <- left_join(dataset2, e5_tobmd, by = "ergoid")
dataset2 <- left_join(dataset2,e5_SES,by="ergoid")
dataset2 <- left_join(dataset2,e5_education,by="ergoid")
dataset2 <- left_join(dataset2, e5_PA, by="ergoid")
dataset2 <- left_join(dataset2, e5_smk, by="ergoid")

calculate_age <- function(dob, enddate) {
  if (is.na(dob) | is.na(enddate)) {
    return(NA)
  } else {
    return(age_calc(as.Date(dob), as.Date(enddate), units = 'years'))
  }
}

dataset2 <- dataset2 %>%
  mutate(
    age = mapply(calculate_age, date_of_birth, dexadate)
  )

# alcohol consumption groups in sexes
dataset2 <- dataset2 %>%
  mutate(
    alc_group = case_when(
      sex == 1 & e5_Alc_tot == 0 ~ 0,
      sex == 1 & e5_Alc_tot > 0 & e5_Alc_tot <= 10 ~ 1,
      sex == 1 & e5_Alc_tot > 10 ~ 2,
      sex == 0 & e5_Alc_tot == 0 ~ 0,
      sex == 0 & e5_Alc_tot > 0 & e5_Alc_tot <= 20 ~ 1,
      sex == 0 & e5_Alc_tot > 20 ~ 2,
      TRUE ~ NA_real_ 
    )
  )

dataset2$nkbmd <- ifelse(!is.na(dataset2$rnkbmd),dataset2$rnkbmd,dataset2$lnkbmd)


# regroup income based on the quantiles
dataset2 <- dataset2 %>% 
  mutate(income_cat1 = case_when(
    income_cat >= 1 & income_cat < 8 ~ 1,
    income_cat >= 8 & income_cat < 10 ~ 2,
    income_cat >= 10 & income_cat < 12 ~ 3,
    income_cat >= 12 ~ 4, 
    income_cat == 77 ~ NA_real_,
    TRUE ~ income_cat)) 

# remove BMD NA & alcohol intake NA
dataset3 <- dataset2 %>%
  filter(!(is.na(lsbmd) | is.na(nkbmd))) # 5819 left

dataset3 <- dataset3 %>%
  filter(!is.na(e5_Alc_dag)) # 5793 left

# remove non-European
dataset3 <- dataset3[dataset3$ancestry == 1,] # 4990 left

# add relatedness
relateness <- relateness_0125[,c(2,4,5,10,7,8,9)]
colnames(relateness)[colnames(relateness) == "IID1"] <- "ergoid"
colnames(relateness)[colnames(relateness) == "IID2"] <- "ergoid2"

relatedness_unique <- relateness %>%
  distinct(ergoid, .keep_all = TRUE)
relatedness_unique$related <- ifelse(relatedness_unique$PI_HAT>=0.125, 1, 0)
write.csv(relatedness_unique, "relatedness_RS_unique.csv")

dataset10 <- left_join(dataset3, relatedness_unique, by="ergoid") 
table(dataset10$related) # 903 related
write.csv(dataset10, "dataset10_includingRelatedness.csv",row.names = F)
```

# dataset11 is the dataset filter out relateness
```{r}
dataset11 <- dataset11 %>% 
  group_by(sex) %>% 
  mutate(std_lsbmd = as.numeric(scale(lsbmd)),
         std_nkbmd = as.numeric(scale(nkbmd)),
         std_totbmd = as.numeric(scale(TOTbmd)),
         std_tsb = as.numeric(scale(lstbs))) %>% 
  ungroup()


dataset11_f <- subset(dataset11, sex == 1)
dataset11_m <- subset(dataset11, sex == 0)
```

# Basic characteristics
```{r}
vars <- c("age","BMI","lsbmd","nkbmd","TOTbmd","lstbs","e5_total_MET","e5_Alc_tot","alc_group","ses_UNESCO_recoded","income_cat1","smk_status")
catvars <- c("alc_group","ses_UNESCO_recoded","income_cat1","smk_status")

tab1 <- CreateTableOne(vars = vars, data = dataset3, factorVars = catvars,strata = "sex")
tab1Mat <- print(tab1, nonnormal = c("e5_total_MET","e5_Alc_tot"),formatOptions = list(big.mark=","), showAllLevels = T)
write.csv(tab1Mat, file = "DescTab_RS.csv")
summary(tab1)
```

## Basic characteristics for Dataset 11
```{r}
# Filtering out relatedness
tab2 <- CreateTableOne(vars = c(vars,"sex"), data = dataset11, factorVars = catvars, strata = "alc_group")
tab2Mat <- print(tab2, nonnormal = c("e5_total_MET","e5_Alc_tot"),formatOptions = list(big.mark=","), showAllLevels = T)
write.csv(tab2Mat, file = "DesTab_RS_withoutRelatedness.csv")
```

# Histogram of alcohol consumption (log-transformed)
```{r}
dataset3$log_alc_dag <- log(dataset3$e5_Alc_dag + 1)

hist_alc_log <- ggplot(dataset3, aes(x = log_alc_dag)) + 
  geom_histogram(binwidth = 0.1, fill = "blue", color = "black") +
  labs(title = "Histogram of log-transformed alcohol consumption", 
       x = "Log-transformed alcohol consumption (drinks/day)", 
       y = "Frequency") +
  theme_minimal()
hist_alc_log
#ggsave("hist_alc_log.png",hist_alc_log, width = 5, height = 4)
```

# Linear Model
## LSBMD
```{r}
ls_mod1 <- lm(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(ls_mod1, conf.int = TRUE)
ls_mod2 <- lm(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(ls_mod2, conf.int = TRUE)
ls_mod3 <- lm(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(ls_mod3, conf.int = TRUE)

ls_mod3_interaction <- lm(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI + log_alc_dag:age, data = dataset11)
tidy(ls_mod3_interaction, conf.int = TRUE)
```

## FNBMD
```{r}
nk_mod1 <- lm(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(nk_mod1, conf.int = T)
nk_mod2 <- lm(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(nk_mod2, conf.int = T)
nk_mod3 <- lm(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(nk_mod3, conf.int = T)
```

## TBBMD
```{r}
TOT_mod1 <- lm(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(TOT_mod1, conf.int = T)
TOT_mod2 <- lm(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(TOT_mod2, conf.int = T)
TOT_mod3 <- lm(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(TOT_mod3, conf.int = T)
```

## LSTBS
```{r}
tbs_mod1 <- lm(std_tbs ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(tbs_mod1, conf.int = T)
tbs_mod2 <- lm(std_tbs ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(tbs_mod2, conf.int = T)
tbs_mod3 <- lm(std_tbs ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(tbs_mod3, conf.int = T)
```

# alcohol-levels analysis in general population
```{r}
dataset11$alc_group <- as.factor(dataset11$alc_group)
```
## LSBMD
```{r}
mod_cat_lsbmd <- lm(scale(lsbmd) ~ alc_group + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(mod_cat_lsbmd, conf.int = TRUE)
```
## FNBMD
```{r}
mod_cat_nkbmd <- lm(scale(nkbmd) ~ alc_group + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(mod_cat_nkbmd, conf.int = TRUE)
```
## TBBMD
```{r}
mod_cat_TOTbmd <- lm(scale(TOTbmd) ~ alc_group + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(mod_cat_TOTbmd, conf.int = TRUE)
```
## LSTBS
```{r}
mod_cat_lstbs <- lm(std_tbs ~ alc_group + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(mod_cat_lstbs, conf.int = TRUE)
```

# Forest plot - general population
```{r}
# LS-NK BMD
Forestplot_LS_NK <- read_excel("Forestplot_RS.xlsx", 
    sheet = "Sheet1")

# Wide format
data_wide <- Forestplot_LS_NK |>
  pivot_wider(names_from = group, values_from = c(mean, lower, upper)) |>
  arrange(labeltext)
tabletext <- data_wide$labeltext
mean  <- as.matrix(data_wide[, c("mean_light", "mean_heavy")])
lower <- as.matrix(data_wide[, c("lower_light", "lower_heavy")])
upper <- as.matrix(data_wide[, c("upper_light", "upper_heavy")])
xticks <- round(pretty(c(lower, upper), n = 5), 2)


png("fp.png", res = 200, width = 1000, height = 800)
Forestplot_LS_NK |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = 0.1, # We set the box size to better visualize the type
             line.margin = 0.3, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE,
             xticks = xticks) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()
```
```{r}
# TO-TBS BMD
Forestplot_TO_TBS <- read_excel("Forestplot.xlsx", 
    sheet = "Sheet2")
fp_2 <- ggplot(Forestplot_TO_TBS,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.2)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=2.0)
fp_2
#ggsave("fp_2.png",width = 6,height = 5)
```

# Sex-stratified analysis - female
# male = 0; female = 1

## Linear model
## LSBMD
```{r}
ls_mod1_f <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_f)
tidy(ls_mod1_f,conf.int = TRUE)
ls_mod2_f <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_f)
tidy(ls_mod2_f,conf.int = TRUE)
ls_mod3_f <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(ls_mod3_f,conf.int = TRUE)
```

## FNBMD
```{r}
ls_mod1_f <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_f)
tidy(ls_mod1_f,conf.int = TRUE)
ls_mod2_f <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_f)
tidy(ls_mod2_f,conf.int = TRUE)
ls_mod3_f <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(ls_mod3_f,conf.int = TRUE)
```

## TBBMD
```{r}
ls_mod1_f <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_f)
tidy(ls_mod1_f)
ls_mod2_f <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_f)
tidy(ls_mod2_f)
ls_mod3_f <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(ls_mod3_f,conf.int = TRUE)
```

## LSTBS
```{r}
ls_mod1_f <- lm(std_tbs ~ log_alc_dag + age + rs_cohort.x, data = dataset11_f)
tidy(ls_mod1_f)
ls_mod2_f <- lm(std_tbs ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_f)
tidy(ls_mod2_f)
ls_mod3_f <- lm(std_tbs ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(ls_mod3_f,conf.int = TRUE)
```

## alcohol-levels analysis - female
## LSBMD
```{r}
mod_cat_lsbmd_f <- lm(std_lsbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(mod_cat_lsbmd_f, conf.int = TRUE)
```
## FNBMD
```{r}
mod_cat_nkbmd_f <- lm(std_nkbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(mod_cat_nkbmd_f, conf.int = TRUE)
```
## TBBMD
```{r}
mod_cat_TOTbmd_f <- lm(std_totbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(mod_cat_TOTbmd_f, conf.int = TRUE)
```
## LSTBS
```{r}
mod_cat_lstbs_f <- lm(std_tbs ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(mod_cat_lstbs_f, conf.int = TRUE)
```

# Forest plot - female
```{r}
# LS-NK BMD
Forestplot_LS_NK_f <- read_excel("Forestplot_F_RS.xlsx",sheet = "Sheet1")
# Wide format
data_wide <- Forestplot_LS_NK_f |>
  pivot_wider(names_from = group, values_from = c(mean, lower, upper)) |>
  arrange(labeltext)
tabletext <- data_wide$labeltext
mean  <- as.matrix(data_wide[, c("mean_light", "mean_heavy")])
lower <- as.matrix(data_wide[, c("lower_light", "lower_heavy")])
upper <- as.matrix(data_wide[, c("upper_light", "upper_heavy")])
xticks <- round(pretty(c(lower, upper), n = 5), 2)


# Codes below should be run in console
png("fp_f.png", width = 1000, height = 800, res = 200)
Forestplot_LS_NK_f |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = .1, # We set the box size to better visualize the type
             line.margin = .3, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE,
             xticks = xticks) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()


# TO-TBS BMD
Forestplot_TO_TBS_f <- read_excel("Forestplot_F.xlsx",sheet = "Sheet2")
fp_2_f <- ggplot(Forestplot_TO_TBS_f,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.5, linewidth = 1)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=3.0)
fp_2_f
#ggsave("fp_2_f.png",width = 6,height = 5)
```

# Sex-stratified analysis - male
## Linear model
## LSBMD
```{r}
ls_mod1_m <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_m)
tidy(ls_mod1_m)
ls_mod2_m <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_m)
tidy(ls_mod2_m)
ls_mod3_m <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(ls_mod3_m,conf.int = TRUE)
```
## FNBMD
```{r}
ls_mod1_m <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_m)
tidy(ls_mod1_m)
ls_mod2_m <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_m)
tidy(ls_mod2_m)
ls_mod3_m <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(ls_mod3_m,conf.int = TRUE)
```
## TBBMD
```{r}
ls_mod1_m <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_m)
tidy(ls_mod1_m)
ls_mod2_m <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_m)
tidy(ls_mod2_m)
ls_mod3_m <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(ls_mod3_m,conf.int = TRUE)
```
## LSTBS
```{r}
ls_mod1_m <- lm(std_tbs ~ log_alc_dag + age + rs_cohort.x, data = dataset11_m)
tidy(ls_mod1_m)
ls_mod2_m <- lm(std_tbs ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_m)
tidy(ls_mod2_m)
ls_mod3_m <- lm(std_tbs ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(ls_mod3_m,conf.int = TRUE)
```
## alcohol-levels analysis - male
## LSBMD
```{r}
mod_cat_lsbmd_m <- lm(std_lsbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(mod_cat_lsbmd_m, conf.int = TRUE)
```
## FNBMD
```{r}
mod_cat_nkbmd_m <- lm(std_nkbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(mod_cat_nkbmd_m, conf.int = TRUE)
```
## TBBMD
```{r}
mod_cat_TOTbmd_m <- lm(std_totbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(mod_cat_TOTbmd_m, conf.int = TRUE)
```
## LSTBS
```{r}
mod_cat_lstbs_m <- lm(std_tbs ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(mod_cat_lstbs_m, conf.int = TRUE)
```
# Forest plot - male
```{r}
# LS-NK BMD
Forestplot_LS_NK_m <- read_excel("Forestplot_M_RS.xlsx", sheet = "Sheet1")

# Wide format
data_wide <- Forestplot_LS_NK_m |>
  pivot_wider(names_from = group, values_from = c(mean, lower, upper)) |>
  arrange(labeltext)
tabletext <- data_wide$labeltext
mean  <- as.matrix(data_wide[, c("mean_light", "mean_heavy")])
lower <- as.matrix(data_wide[, c("lower_light", "lower_heavy")])
upper <- as.matrix(data_wide[, c("upper_light", "upper_heavy")])
xticks <- round(pretty(c(lower, upper), n = 5), 2)

png("fp_m.png", width = 1000, height = 800,res = 200)
Forestplot_LS_NK_m |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = .1, # We set the box size to better visualize the type
             line.margin = 0.3, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE,
             xticks = xticks) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()


# TO-TBS BMD
Forestplot_TO_TBS_m <- read_excel("Forestplot_M.xlsx", sheet = "Sheet2")
fp_2_m <- ggplot(Forestplot_TO_TBS_m,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.5,linewidth =1)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=3.0)
fp_2_m
#ggsave("fp_2_m.png",width = 6,height = 5)
```

# Age-stratified
```{r}
dataset11_young <- subset(dataset11, age<60)
dataset11_old <- subset(dataset11, age>=60)
```
## Age < 60yrs
## LSBMD
```{r}
ls_mod3_young <- lm(std_lsbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_young)
tidy(ls_mod3_young, conf.int = TRUE)
```
## FNBMD
```{r}
nk_mod3_young <- lm(std_nkbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_young)
tidy(nk_mod3_young, conf.int = TRUE)
```
## TBBMD
```{r}
tot_mod3_young <- lm(std_totbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_young)
tidy(tot_mod3_young, conf.int = TRUE)
```
## LSTBS
```{r}
tbs_mod3_young <- lm(std_tbs ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_young)
tidy(tbs_mod3_young, conf.int = TRUE)
```
## Age >= 60yrs
## LSBMD
```{r}
ls_mod3_old <- lm(std_lsbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_old)
tidy(ls_mod3_old, conf.int = TRUE)
```
## FNBMD
```{r}
nk_mod3_old <- lm(std_nkbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_old)
tidy(nk_mod3_old, conf.int = TRUE)
```
## TBBMD
```{r}
tot_mod3_old <- lm(std_totbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_old)
tidy(tot_mod3_old, conf.int = TRUE)
```
## LSTBS
```{r}
tbs_mod3_old <- lm(std_tbs ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_old)
tidy(tbs_mod3_old, conf.int = TRUE)
```
# 1-sample MR for Alcohol consumption
## Input PGS files for alcohol consumption
```{r}
PGS_71_rs1 <- read.table("PGS-SNPs_r0.001_10000kb.csv-RS1_72.txt",header = T)
PGS_71_rs2 <- read.table("PGS-SNPs_r0.001_10000kb.csv-RS2_72.txt",header = T)
PGS_71_rs3 <- read.table("PGS-SNPs_r0.001_10000kb.csv-RS3_72.txt",header = T)
PGS_71_rs <- rbind(PGS_71_rs1,PGS_71_rs2)
PGS_71_rs <- rbind(PGS_71_rs,PGS_71_rs3)

PGS_71_rs <- PGS_71_rs %>% rownames_to_column(var = "ergoid")
PGS_71_rs <- PGS_71_rs %>% rename(PGS_71 = PGS, wPGS_71 = wPGS)
PGS_71_rs$ergoid <- as.numeric(PGS_71_rs$ergoid)

# Add PGS into dataset
dataset12 <- left_join(dataset11, PGS_71_rs, by = "ergoid")
```

# Histogram of PGS
```{r}
hist_PGS <- ggplot(dataset12, aes(x = PGS_71_rs)) + 
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Histogram of alcohol consumption PGS in the Rotterdam study", 
       x = "Alcohol consumption PGS", 
       y = "Frequency") +
  theme_minimal()
hist_PGS
#ggsave("hist_PGS.png", plot = hist_PGS, width = 6, height = 5)
```

## 1-sample MR process
## LSBMD
```{r}
iv_ls_71 <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_71 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_ls_71, conf.int = T)
```
## FNBMD
```{r}
iv_nk_71 <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_71 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_nk_71, conf.int = T)
```
## TBBMD
```{r}
iv_TOT_71 <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_71 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_TOT_71, conf.int = T)
```
## LSTBS
```{r}
iv_tbs_71 <- ivreg(std_tbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_71 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_tbs_71, conf.int = T)
```


# 1-sample MR for PAU
## Input PGS files for PAU
```{r}
# PRS for PAU - r0.001
PRS_PAU_r0001_RS1 <- read.table("PGS-clumped_PAU_r0001.csv-RS1.txt",header = TRUE)
PRS_PAU_r0001_RS2 <- read.table("PGS-clumped_PAU_r0001.csv-RS2.txt",header = TRUE)
PRS_PAU_r0001_RS3 <- read.table("PGS-clumped_PAU_r0001.csv-RS3.txt",header = TRUE)

PRS_PAU_r0001_RS <- bind_rows(list(PRS_PAU_r0001_RS1, PRS_PAU_r0001_RS2, PRS_PAU_r0001_RS3))
PRS_PAU_r0001_RS <- PRS_PAU_r0001_RS %>%
  rownames_to_column(var = "ergoid") %>%
  rename(PGS_PAU_r0001 = PGS,
         wPGS_PAU_r0001 = wPGS)

PRS_PAU_r0001_RS$ergoid <- as.numeric(PRS_PAU_r0001_RS$ergoid)
dataset_with_PAUPRS <- left_join(dataset_with_PAUPRS, PRS_PAU_r0001_RS, by = "ergoid")
```

# 1-sample MR - PAU - r2 0.001 process
## LSBMD
```{r}
iv_ls_PAU_r0001 <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU_r0001 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_ls_PAU_r0001, conf.int = TRUE)
```
## FNBMD
```{r}
iv_nk_PAU_r0001 <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU_r0001 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_nk_PAU_r0001, conf.int = TRUE)
```
## TBBMD
```{r}
iv_TOT_PAU_r0001 <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU_r0001 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_TOT_PAU_r0001, conf.int = TRUE)
```
## LSTBS
```{r}
iv_tbs_PAU_r0001 <- ivreg(std_lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU_r0001 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_tbs_PAU_r0001, conf.int = TRUE)
```

# 1-sample MR SNPs without UKB
## Input PGS for alcohol consumption without UKB
```{r}
PRS_withoutUKB_RS1_71 <- read.table("PGS-SNPs_r0.001_10000kb_withoutUKB.csv-RS1withoutUKB_72.txt", header = TRUE)
PRS_withoutUKB_RS2_71 <- read.table("PGS-SNPs_r0.001_10000kb_withoutUKB.csv-RS2withoutUKB_72.txt", header = TRUE)
PRS_withoutUKB_RS3_71 <- read.table("PGS-SNPs_r0.001_10000kb_withoutUKB.csv-RS3withoutUKB_72.txt", header = TRUE)

PRS_withoutUKB_RS_71 <- bind_rows(list(PRS_withoutUKB_RS1_71, PRS_withoutUKB_RS2_71, PRS_withoutUKB_RS3_71))
PRS_withoutUKB_RS_71 <- PRS_withoutUKB_RS_71 %>%
  rownames_to_column(var = "ergoid") %>%
  rename(PGS_withoutUKB_71 = PGS,
         wPGS_withoutUKB_71 = wPGS)
PRS_withoutUKB_RS_71$ergoid <- as.numeric(PRS_withoutUKB_RS_71$ergoid)

dataset_PRS_withoutUKB <- left_join(dataset_PRS_withoutUKB, PRS_withoutUKB_RS_71, by = "ergoid")
```

# 1-sample MR without UKB process
## LSBMD
```{r}
iv_ls_withoutUKB_71 <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB_71 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_ls_withoutUKB_71, conf.int = TRUE)
```
## FNBMD
```{r}
iv_nk_withoutUKB_71 <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB_71 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_nk_withoutUKB_71, conf.int = TRUE)
```
## TBBMD
```{r}
iv_TOT_withoutUKB_71 <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB_71 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_TOT_withoutUKB_71, conf.int = TRUE)
```
## LSTBS
```{r}
iv_tbs_withoutUKB_71 <- ivreg(std_tbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB_71 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_tbs_withoutUKB_71, conf.int = TRUE)
```

