---
title: "RS_ergo5"
author: "Xinpeng"
date: "2024-06-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(root.dir = "//store/department/im/share/personal/059909/PHD/Bone & Alcohol/ERGO5/RS_E5")
```


# Settings
```{r}
#setwd("//store/department/im/share/personal/059909/PHD/Bone & Alcohol/ERGO5/RS_E5")

# Library packages
library(dplyr)
library(ggplot2)
library(haven)
library(eeptools)
library(sjlabelled)
library(readxl)
library(tidyr)
library(AER)
library(tableone)
library(Hmisc)
library(broom)
library(forestplot)
library(tidyverse)
```


# Input original datasets
```{r}
RoterdamStudy_Basics2014 <- read_sav("RoterdamStudy_Basics2014.sav")
RS_GeneticEtnicity_11_JAN_2016_ <- read_sav("RS_GeneticEtnicity_(11-JAN-2016).sav")
e5_intvw_Alcoholperday_05_11_2019 <- read_sav("e5_intvw_Alcoholperday_05-11-2019.sav")
e5_intvw_SMOKING_22_mar_2021_ <- read_sav("e5_intvw_SMOKING_(22-mar-2021).sav")
Ergo5_RSI_5_RSII_3_RSIII_2_Femur_DXA_iDXA_BMD <- read_sav("Ergo5_(RSI-5_RSII-3_RSIII-2)_Femur_DXA_iDXA_BMD.sav")
e5_5_ls_bmd_ls_tbs_15_feb_2022_ <- read_sav("e5_(5)_ls.bmd ls.tbs_(15-feb-2022).sav")
RS_GeneticEtnicity_11_JAN_2016_ <- read_sav("RS_GeneticEtnicity_(11-JAN-2016).sav")
e5_intvw_SES_07_feb_2014_ <- read_sav("e5_intvw_SES_(07-feb-2014).sav")
e_InformedConsent_24_MAY_2018_ <- read_sav("e_InformedConsent_(24-MAY-2018).sav")
Ergo5_RSI_5_RSII_3_RSIII_2_Total_Body_DXA_iDXA_BMD <- read_sav("Ergo5_(RSI-5_RSII-3_RSIII-2)_Total_Body_DXA_iDXA_BMD.sav")
e1epej_SES <- read_sav("e1epej_SES.sav")
Education_RS_I_II_III_UNESCO_class_12_MAR_2015_ <- read_sav("Education RS-I-II-III (UNESCO class)_(12-MAR-2015).sav")
e5_5_ANTHROPO_10_dec_2015_ <- read_sav("e5_(5)_ANTHROPO_(10-dec-2015).sav")
ERGO_5_RS_I_5_II_3_III_2_LASA_07_2020_ <- read_sav("ERGO 5 (RS-I-5, II-3, III-2) - LASA (07-2020).sav")
relateness_0125 <- read.table("//store/department/im/share/projects/Fernando_Group/Alcohol_BMD_Xinpeng/Rotterdam Study/related_0125.txt", header = T)
```

# Create datasets for analysis
```{r}
Basic_ergo5 <- RoterdamStudy_Basics2014
Basic_ergo5 <- left_join(Basic_ergo5,RS_GeneticEtnicity_11_JAN_2016_,by="ergoid")

e5_InformConsent <- e_InformedConsent_24_MAY_2018_[,c(1,32)]
e5_InformConsent <- e5_InformConsent %>%
  rename(e5_informconsent = e5_4109) %>%
  mutate(e5_informconsent = na_if(e5_informconsent, 8)) %>%
  mutate(e5_informconsent = na_if(e5_informconsent, 9)) %>%
  mutate(e5_informconsent = case_when(
    e5_informconsent == 1 ~ 1,
    e5_informconsent == 2 ~ 1,
    e5_informconsent == 4 ~ 0,
    TRUE ~ e5_informconsent
  ))

dataset1 <- left_join(Basic_ergo5,e5_InformConsent, by ="ergoid")
dataset1 <- dataset1[!is.na(dataset1$e5_informconsent),]
dataset1 <- dataset1[dataset1$e5_informconsent == 1,] # in total 7146 participants have inform consent

# outcomes & exposures & covariates
e5_alc <- e5_intvw_Alcoholperday_05_11_2019[,c(1,21:24)]
e5_ls_bmd_tbs <- e5_5_ls_bmd_ls_tbs_15_feb_2022_[,c(1,6,13,14)]
e5_fnbmd <- Ergo5_RSI_5_RSII_3_RSIII_2_Femur_DXA_iDXA_BMD[,c(1,3,6,9)]
e5_tobmd <- Ergo5_RSI_5_RSII_3_RSIII_2_Total_Body_DXA_iDXA_BMD[,c(1,6)]
e5_SES <- e1epej_SES[,c(1,6)]
e5_smk <- e5_intvw_SMOKING_22_mar_2021_[,c(1,7)]
e5_education <- Education_RS_I_II_III_UNESCO_class_12_MAR_2015_[,c(1,3)]
e5_PA <- ERGO_5_RS_I_5_II_3_III_2_LASA_07_2020_[,c(1,3,4)]
e5_BMI <- e5_5_ANTHROPO_10_dec_2015_[,c(1,6,7)] %>%
  mutate(BMI = e5_230/(e5_229/100)^2)

e5_smk <- e5_intvw_SMOKING_22_mar_2021_ %>%
  mutate(smk_status = case_when(
    e5_EILF6 == 0 & e5_EILFE==0 & e5_EILF4==0 & e5_EILF5==0 ~ 0,
    (e5_EILF4==1|e5_EILF4==2|e5_EILF4==3)|e5_EILF5==1|(e5_EILF6==0&e5_EILFE==1) ~ 1,
    (e5_EILF4==4|e5_EILF4==5|e5_EILF4==6)|e5_EILF5==2|e5_EILF5==3|e5_EILF6==1 ~ 2
  )) %>%
  select(ergoid, smk_status)
  

dataset2 <- left_join(dataset1, e5_BMI, by = "ergoid")
dataset2 <- left_join(dataset2, e5_alc, by = "ergoid")
dataset2 <- left_join(dataset2, e5_ls_bmd_tbs, by = "ergoid")
dataset2 <- left_join(dataset2, e5_fnbmd, by = "ergoid")
dataset2 <- left_join(dataset2, e5_tobmd, by = "ergoid")
dataset2 <- left_join(dataset2,e5_SES,by="ergoid")
dataset2 <- left_join(dataset2,e5_education,by="ergoid")
dataset2 <- left_join(dataset2, e5_PA, by="ergoid")
dataset2 <- left_join(dataset2, e5_smk, by="ergoid")

calculate_age <- function(dob, enddate) {
  if (is.na(dob) | is.na(enddate)) {
    return(NA)
  } else {
    return(age_calc(as.Date(dob), as.Date(enddate), units = 'years'))
  }
}

dataset2 <- dataset2 %>%
  mutate(
    age = mapply(calculate_age, date_of_birth, dexadate)
  )

# alcohol consumption groups in sexes
dataset2 <- dataset2 %>%
  mutate(
    alc_group = case_when(
      sex == 1 & e5_Alc_tot == 0 ~ 0,
      sex == 1 & e5_Alc_tot > 0 & e5_Alc_tot <= 10 ~ 1,
      sex == 1 & e5_Alc_tot > 10 ~ 2,
      sex == 0 & e5_Alc_tot == 0 ~ 0,
      sex == 0 & e5_Alc_tot > 0 & e5_Alc_tot <= 20 ~ 1,
      sex == 0 & e5_Alc_tot > 20 ~ 2,
      TRUE ~ NA_real_ 
    )
  )

dataset2$nkbmd <- ifelse(!is.na(dataset2$rnkbmd),dataset2$rnkbmd,dataset2$lnkbmd)


# regroup income based on the quantiles
dataset2 <- dataset2 %>% 
  mutate(income_cat1 = case_when(
    income_cat >= 1 & income_cat < 8 ~ 1,
    income_cat >= 8 & income_cat < 10 ~ 2,
    income_cat >= 10 & income_cat < 12 ~ 3,
    income_cat >= 12 ~ 4, 
    income_cat == 77 ~ NA_real_,
    TRUE ~ income_cat)) 

# remove BMD NA & alcohol intake NA
dataset3 <- dataset2 %>%
  filter(!(is.na(lsbmd) | is.na(nkbmd))) # 5819 left

dataset3 <- dataset3 %>%
  filter(!is.na(e5_Alc_dag)) # 5793 left

# remove non-European
dataset3 <- dataset3[dataset3$ancestry == 1,] # 4990 left

# add relatedness
relateness <- relateness_0125[,c(2,4,5,10,7,8,9)]
colnames(relateness)[colnames(relateness) == "IID1"] <- "ergoid"
colnames(relateness)[colnames(relateness) == "IID2"] <- "ergoid2"

relatedness_unique <- relateness %>%
  distinct(ergoid, .keep_all = TRUE)
relatedness_unique$related <- ifelse(relatedness_unique$PI_HAT>=0.125, 1, 0)
write.csv(relatedness_unique, "relatedness_RS_unique.csv")

dataset10 <- left_join(dataset3, relatedness_unique, by="ergoid") 
table(dataset10$related) # 903 related
write.csv(dataset10, "dataset10_includingRelatedness.csv",row.names = F)
```

# dataset11 is the dataset filter out relateness
```{r}
dataset11 <- dataset11 |>
  group_by(sex) |>
  mutate(std_lsbmd = as.vector(scale(lsbmd)),
         std_nkbmd = as.vector(scale(nkbmd)),
         std_totbmd = as.vector(scale(TOTbmd))) |>
  ungroup()


dataset11_f <- subset(dataset11, sex == 1)
dataset11_m <- subset(dataset11, sex == 0)
```

# Basic characteristics
```{r}
vars <- c("age","BMI","lsbmd","nkbmd","TOTbmd","lstbs","e5_total_MET","e5_Alc_tot","alc_group","ses_UNESCO_recoded","income_cat1","smk_status")
catvars <- c("alc_group","ses_UNESCO_recoded","income_cat1","smk_status")

tab1 <- CreateTableOne(vars = vars, data = dataset3, factorVars = catvars,strata = "sex")
tab1Mat <- print(tab1, nonnormal = c("e5_total_MET","e5_Alc_tot"),formatOptions = list(big.mark=","), showAllLevels = T)
write.csv(tab1Mat, file = "DescTab_RS.csv")
summary(tab1)
```

```{r}
# Filtering out relatedness
tab2 <- CreateTableOne(vars = c(vars,"sex"), data = dataset11, factorVars = catvars, strata = "alc_group")
tab2Mat <- print(tab2, nonnormal = c("e5_total_MET","e5_Alc_tot"),formatOptions = list(big.mark=","), showAllLevels = T)
write.csv(tab2Mat, file = "DesTab_RS_withoutRelatedness.csv")
```
# Histogram of alcohol consumption
```{r}
# 1 drink = 10 g alcohol
hist_alc_tot <- ggplot(dataset3, aes(x = e5_Alc_tot)) + 
  geom_histogram(binwidth = 10, fill = "blue", color = "black") +
  labs(title = "Histogram of alcohol consumption", 
       x = "Alcohol consumption (g/day)", 
       y = "Frequency") +
  theme_minimal()
hist_alc_tot
#ggsave("hist_alc_tot.png",hist_alc_tot, width = 5, height = 4)
```

```{r}
dataset3$log_alc_dag <- log(dataset3$e5_Alc_dag + 1)

hist_alc_log <- ggplot(dataset3, aes(x = log_alc_dag)) + 
  geom_histogram(binwidth = 0.1, fill = "blue", color = "black") +
  labs(title = "Histogram of log-transformed alcohol consumption", 
       x = "Log-transformed alcohol consumption (drinks/day)", 
       y = "Frequency") +
  theme_minimal()
hist_alc_log
#ggsave("hist_alc_log.png",hist_alc_log, width = 5, height = 4)
```


# Linear Model
```{r}
# LSBMD
dataset3$sex <- as.factor(dataset3$sex)
dataset3$rs_cohort.x <- as.factor(dataset3$rs_cohort.x)
dataset3$income_cat1 <- as.factor(dataset3$income_cat1)
dataset3$ses_UNESCO_recoded <- as.factor(dataset3$ses_UNESCO_recoded)
dataset3$smk_status <- as.factor(dataset3$smk_status)

dataset3 <- dataset3 |>
   group_by(sex) |>
   mutate(std_lsbmd = as.vector(scale(lsbmd))) |>
   ungroup()

dataset11 <- dataset11 %>%
  group_by(sex) %>% 
  mutate(std_tbs = as.numeric(scale(lstbs))) %>% 
  ungroup()

ls_mod1 <- lm(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(ls_mod1, conf.int = TRUE)
ls_mod2 <- lm(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(ls_mod2, conf.int = TRUE)
ls_mod3 <- lm(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(ls_mod3, conf.int = TRUE)

ls_mod3_interaction <- lm(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI + log_alc_dag:age, data = dataset11)
tidy(ls_mod3_interaction, conf.int = TRUE)
```

```{r}
# NKBMD
nk_mod1 <- lm(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(nk_mod1, conf.int = T)
nk_mod2 <- lm(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(nk_mod2, conf.int = T)
nk_mod3 <- lm(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(nk_mod3, conf.int = T)

```
```{r}
# TOBMD
TOT_mod1 <- lm(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(TOT_mod1, conf.int = T)
TOT_mod2 <- lm(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(TOT_mod2, conf.int = T)
TOT_mod3 <- lm(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(TOT_mod3, conf.int = T)
```

```{r}
# LSTBS
tbs_mod1 <- lm(lstbs ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(tbs_mod1, conf.int = T)
tbs_mod2 <- lm(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(tbs_mod2, conf.int = T)
tbs_mod3 <- lm(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(tbs_mod3, conf.int = T)
```

```{r}
tbs_mod1 <- lm(std_tbs ~ log_alc_dag + age + sex + rs_cohort.x, data = dataset11)
tidy(tbs_mod1, conf.int = T)
tbs_mod2 <- lm(std_tbs ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11)
tidy(tbs_mod2, conf.int = T)
tbs_mod3 <- lm(std_tbs ~ log_alc_dag + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
tidy(tbs_mod3, conf.int = T)
```

## alcohol-levels analysis in general population
```{r}
dataset11$alc_group <- as.factor(dataset11$alc_group)

mod_cat_lsbmd <- lm(scale(lsbmd) ~ alc_group + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)


mod_cat_nkbmd <- lm(scale(nkbmd) ~ alc_group + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)


mod_cat_TOTbmd <- lm(scale(TOTbmd) ~ alc_group + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)


mod_cat_lstbs <- lm(std_tbs ~ alc_group + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)

tidy(mod_cat_lsbmd, conf.int = TRUE)
tidy(mod_cat_nkbmd, conf.int = TRUE)
tidy(mod_cat_TOTbmd, conf.int = TRUE)
tidy(mod_cat_lstbs, conf.int = TRUE)
```
## trend analysis - general population
```{r}
mod_trend_lsbmd <- lm(scale(lsbmd) ~ as.numeric(alc_group) + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
summary(mod_trend_lsbmd)
mod_trend_nkbmd <- lm(scale(nkbmd) ~ as.numeric(alc_group) + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
summary(mod_trend_nkbmd)
mod_trend_TOTbmd <- lm(scale(TOTbmd) ~ as.numeric(alc_group) + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
summary(mod_trend_TOTbmd)
mod_trend_lstbs <- lm(lstbs ~ as.numeric(alc_group) + age + sex + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11)
summary(mod_trend_lstbs)
```
## Forest plot - general population
```{r}
# LS-NK BMD
Forestplot_LS_NK <- read_excel("Forestplot_RS.xlsx", 
    sheet = "Sheet1")

# Wide format
data_wide <- Forestplot_LS_NK |>
  pivot_wider(names_from = group, values_from = c(mean, lower, upper)) |>
  arrange(labeltext)
tabletext <- data_wide$labeltext
mean  <- as.matrix(data_wide[, c("mean_light", "mean_heavy")])
lower <- as.matrix(data_wide[, c("lower_light", "lower_heavy")])
upper <- as.matrix(data_wide[, c("upper_light", "upper_heavy")])
xticks <- round(pretty(c(lower, upper), n = 5), 2)


png("fp.png", res = 200, width = 1000, height = 800)
Forestplot_LS_NK |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = 0.1, # We set the box size to better visualize the type
             line.margin = 0.3, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE,
             xticks = xticks) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()
```
```{r}
# TO-TBS BMD
Forestplot_TO_TBS <- read_excel("Forestplot.xlsx", 
    sheet = "Sheet2")
fp_2 <- ggplot(Forestplot_TO_TBS,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.2)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=2.0)
fp_2
#ggsave("fp_2.png",width = 6,height = 5)
```

# Sex-stratified analysis - female
```{r}
# LSBMD
# male = 0; female = 1
dataset3$std_lsbmd <- scale(dataset3$lsbmd)
dataset3$std_nkbmd <- scale(dataset3$nkbmd)
dataset3$std_totbmd <- scale(dataset3$TOTbmd)

dataset3 <- dataset3 |>
  group_by(sex) |>
  mutate(std_lsbmd = as.vector(scale(lsbmd))) |>
  ungroup()

dataset3_f <- subset(dataset3, sex == 1)
dataset3_m <- subset(dataset3, sex == 0)

dataset11_f <- subset(dataset11, sex == 1)
dataset11_m <- subset(dataset11, sex == 0)

ls_mod1_f <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_f)
tidy(ls_mod1_f,conf.int = TRUE)
ls_mod2_f <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_f)
tidy(ls_mod2_f,conf.int = TRUE)
ls_mod3_f <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(ls_mod3_f,conf.int = TRUE)
```

```{r}
# NKBMD
ls_mod1_f <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_f)
tidy(ls_mod1_f,conf.int = TRUE)
ls_mod2_f <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_f)
tidy(ls_mod2_f,conf.int = TRUE)
ls_mod3_f <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(ls_mod3_f,conf.int = TRUE)

```
```{r}
# TOBMD
ls_mod1_f <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_f)
tidy(ls_mod1_f)
ls_mod2_f <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_f)
tidy(ls_mod2_f)
ls_mod3_f <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(ls_mod3_f,conf.int = TRUE)

```
```{r}
# LSTBS
ls_mod1_f <- lm(lstbs ~ log_alc_dag + age + rs_cohort.x, data = dataset11_f)
tidy(ls_mod1_f)
ls_mod2_f <- lm(lstbs ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_f)
tidy(ls_mod2_f)
ls_mod3_f <- lm(lstbs ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(ls_mod3_f,conf.int = TRUE)
```


## alcohol-levels analysis - female
```{r}
mod_cat_lsbmd_f <- lm(std_lsbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(mod_cat_lsbmd_f, conf.int = TRUE)
mod_cat_nkbmd_f <- lm(std_nkbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(mod_cat_nkbmd_f, conf.int = TRUE)
mod_cat_TOTbmd_f <- lm(std_totbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(mod_cat_TOTbmd_f, conf.int = TRUE)
mod_cat_lstbs_f <- lm(std_tbs ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_f)
tidy(mod_cat_lstbs_f, conf.int = TRUE)


```

## trend analysis - female
```{r}
mod_trend_lsbmd_f <- lm(std_lsbmd ~ as.numeric(alc_group) + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset3_f)
summary(mod_trend_lsbmd_f)
mod_trend_nkbmd_f <- lm(std_nkbmd ~ as.numeric(alc_group) + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset3_f)
summary(mod_trend_nkbmd_f)
mod_trend_TOTbmd_f <- lm(std_totbmd ~ as.numeric(alc_group) + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset3_f)
summary(mod_trend_TOTbmd_f)
mod_trend_lstbs_f <- lm(lstbs ~ as.numeric(alc_group) + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset3_f)
summary(mod_trend_lstbs_f)
```

## Forest plot - female
```{r}
# LS-NK BMD
Forestplot_LS_NK_f <- read_excel("Forestplot_F_RS.xlsx",sheet = "Sheet1")
# Wide format
data_wide <- Forestplot_LS_NK_f |>
  pivot_wider(names_from = group, values_from = c(mean, lower, upper)) |>
  arrange(labeltext)
tabletext <- data_wide$labeltext
mean  <- as.matrix(data_wide[, c("mean_light", "mean_heavy")])
lower <- as.matrix(data_wide[, c("lower_light", "lower_heavy")])
upper <- as.matrix(data_wide[, c("upper_light", "upper_heavy")])
xticks <- round(pretty(c(lower, upper), n = 5), 2)


# Codes below should be run in console
png("fp_f.png", width = 1000, height = 800, res = 200)
Forestplot_LS_NK_f |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = .1, # We set the box size to better visualize the type
             line.margin = .3, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE,
             xticks = xticks) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()


# TO-TBS BMD
Forestplot_TO_TBS_f <- read_excel("Forestplot_F.xlsx",sheet = "Sheet2")
fp_2_f <- ggplot(Forestplot_TO_TBS_f,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.5, linewidth = 1)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=3.0)
fp_2_f
#ggsave("fp_2_f.png",width = 6,height = 5)
```

# Sex-stratified analysis - male
```{r}
# LSBMD
ls_mod1_m <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_m)
tidy(ls_mod1_m)
ls_mod2_m <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_m)
tidy(ls_mod2_m)
ls_mod3_m <- lm(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(ls_mod3_m,conf.int = TRUE)
```

```{r}
# NKBMD
ls_mod1_m <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_m)
tidy(ls_mod1_m)
ls_mod2_m <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_m)
tidy(ls_mod2_m)
ls_mod3_m <- lm(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(ls_mod3_m,conf.int = TRUE)
```

```{r}
# TOBMD
ls_mod1_m <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x, data = dataset11_m)
tidy(ls_mod1_m)
ls_mod2_m <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_m)
tidy(ls_mod2_m)
ls_mod3_m <- lm(std_totbmd ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(ls_mod3_m,conf.int = TRUE)
```

```{r}
# LSTBS
ls_mod1_m <- lm(lstbs ~ log_alc_dag + age + rs_cohort.x, data = dataset11_m)
tidy(ls_mod1_m)
ls_mod2_m <- lm(lstbs ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET, data = dataset11_m)
tidy(ls_mod2_m)
ls_mod3_m <- lm(lstbs ~ log_alc_dag + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(ls_mod3_m,conf.int = TRUE)

```

## alcohol-levels analysis - male
```{r}
mod_cat_lsbmd_m <- lm(std_lsbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(mod_cat_lsbmd_m, conf.int = TRUE)
mod_cat_nkbmd_m <- lm(std_nkbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(mod_cat_nkbmd_m, conf.int = TRUE)
mod_cat_TOTbmd_m <- lm(std_totbmd ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(mod_cat_TOTbmd_m, conf.int = TRUE)
mod_cat_lstbs_m <- lm(std_tbs ~ alc_group + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
tidy(mod_cat_lstbs_m, conf.int = TRUE)
```

## trend analysis - male
```{r}
mod_trend_lsbmd_m <- lm(std_lsbmd ~ as.numeric(alc_group) + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
summary(mod_trend_lsbmd_m)
mod_trend_nkbmd_m <- lm(std_nkbmd ~ as.numeric(alc_group) + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
summary(mod_trend_nkbmd_m)
mod_trend_TOTbmd_m <- lm(std_totbmd ~ as.numeric(alc_group) + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
summary(mod_trend_TOTbmd_m)
mod_trend_lstbs_m <- lm(lstbs ~ as.numeric(alc_group) + age + rs_cohort.x + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_m)
summary(mod_trend_lstbs_m)
```

## Forest plot - male
```{r}
# LS-NK BMD
Forestplot_LS_NK_m <- read_excel("Forestplot_M_RS.xlsx", sheet = "Sheet1")

# Wide format
data_wide <- Forestplot_LS_NK_m |>
  pivot_wider(names_from = group, values_from = c(mean, lower, upper)) |>
  arrange(labeltext)
tabletext <- data_wide$labeltext
mean  <- as.matrix(data_wide[, c("mean_light", "mean_heavy")])
lower <- as.matrix(data_wide[, c("lower_light", "lower_heavy")])
upper <- as.matrix(data_wide[, c("upper_light", "upper_heavy")])
xticks <- round(pretty(c(lower, upper), n = 5), 2)

png("fp_m.png", width = 1000, height = 800,res = 200)
Forestplot_LS_NK_m |> 
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = .1, # We set the box size to better visualize the type
             line.margin = 0.3, # We need to add this to avoid crowding
             xlab = "Beta (95% CI)",
             vertices = TRUE,
             xticks = xticks) |>
  fp_add_lines() |>
  fp_set_style(box = c("blue", "red") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE),
               lines = "darkblue",
               txt_gp = fpTxtGp(ticks = gpar(fontfamily = "", cex = 0.8),
                                xlab = gpar(fontfamily = "", cex = 0.8))) |>
  fp_set_zebra_style("#F5F9F9")
dev.off()


# TO-TBS BMD
Forestplot_TO_TBS_m <- read_excel("Forestplot_M.xlsx", sheet = "Sheet2")
fp_2_m <- ggplot(Forestplot_TO_TBS_m,aes(beta,phenotype,col=phenotype)) +
  geom_errorbarh(aes(xmax=Upper, xmin=Lower),height=0.5,linewidth =1)+
  geom_vline(aes(xintercept=0)) +
  geom_point(size=3.0)
fp_2_m
#ggsave("fp_2_m.png",width = 6,height = 5)
```
# Age-stratified
````{r}
dataset11_young <- subset(dataset11, age<60)
dataset11_old <- subset(dataset11, age>=60)
```

## Age<60yrs
```{r}
ls_mod3_young <- lm(std_lsbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_young)
tidy(ls_mod3_young, conf.int = TRUE)
```
```{r}
nk_mod3_young <- lm(std_nkbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_young)
tidy(nk_mod3_young, conf.int = TRUE)
```
```{r}
tot_mod3_young <- lm(std_totbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_young)
tidy(tot_mod3_young, conf.int = TRUE)
```
```{r}
tbs_mod3_young <- lm(lstbs ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_young)
tidy(tbs_mod3_young, conf.int = TRUE)
```
## Age>=60yrs
```{r}
ls_mod3_old <- lm(std_lsbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_old)
tidy(ls_mod3_old, conf.int = TRUE)
```
```{r}
nk_mod3_old <- lm(std_nkbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_old)
tidy(nk_mod3_old, conf.int = TRUE)
```
```{r}
tot_mod3_old <- lm(std_totbmd ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_old)
tidy(tot_mod3_old, conf.int = TRUE)
```
```{r}
tbs_mod3_old <- lm(lstbs ~ log_alc_dag + age + sex + income_cat1 + ses_UNESCO_recoded + smk_status + e5_total_MET + BMI, data = dataset11_old)
tidy(tbs_mod3_old, conf.int = TRUE)
```

# 1-sample MR
```{r}
# input PGS
PGS_rs1 <- read.table("PGS-ALC_PRS_sum_withsnplist_matchInbothcohorts.csv-RS1.txt",header = T,)
PGS_rs2 <- read.table("PGS-ALC_PRS_sum_withsnplist_matchInbothcohorts.csv-RS2.txt",header = T)
PGS_rs3 <- read.table("PGS-ALC_PRS_sum_withsnplist_matchInbothcohorts.csv-RS3.txt",header = T)
PGS_rs <- rbind(PGS_rs1,PGS_rs2)
PGS_rs <- rbind(PGS_rs,PGS_rs3)



PGS_77_rs1 <- read.table("PGS-ALC_PRS_sum_withsnplist_matchInbothcohorts_77.csv-RS1_77.txt",header = T,)
PGS_77_rs2 <- read.table("PGS-ALC_PRS_sum_withsnplist_matchInbothcohorts_77.csv-RS2_77.txt",header = T)
PGS_77_rs3 <- read.table("PGS-ALC_PRS_sum_withsnplist_matchInbothcohorts_77.csv-RS3_77.txt",header = T)
PGS_77_rs <- rbind(PGS_77_rs1,PGS_77_rs2)
PGS_77_rs <- rbind(PGS_77_rs,PGS_77_rs3)

PGS_77_rs <- PGS_77_rs %>% rownames_to_column(var = "ergoid")
PGS_77_rs <- PGS_77_rs %>% rename(PGS_77 = PGS, wPGS_77 = wPGS)
PGS_77_rs$ergoid <- as.numeric(PGS_77_rs$ergoid)

PRS_RS1_USEPROXY <- read.table("PGS-ALC-RS1_useProxy_96.txt", header = TRUE)
PRS_RS2_USEPROXY <- read.table("PGS-ALC-RS2_useProxy_96.txt", header = TRUE)
PRS_RS3_USEPROXY <- read.table("PGS-ALC-RS3_useProxy_96.txt", header = TRUE)
PRS_RS1_USEPROXY <- rownames_to_column(PRS_RS1_USEPROXY,var = "ergoid")
PRS_RS2_USEPROXY <- rownames_to_column(PRS_RS2_USEPROXY,var = "ergoid")
PRS_RS3_USEPROXY <- rownames_to_column(PRS_RS3_USEPROXY,var = "ergoid")

PRS_RS_USEPROXY <- rbind(PRS_RS1_USEPROXY, PRS_RS2_USEPROXY)
PRS_RS_USEPROXY <- rbind(PRS_RS_USEPROXY, PRS_RS3_USEPROXY)
PRS_RS_USEPROXY <- PRS_RS_USEPROXY %>%
  rename(PGS_useProxy = PGS, wPGS_useProxy = wPGS)
PRS_RS_USEPROXY$ergoid <- as.numeric(PRS_RS_USEPROXY$ergoid)

dataset15 <- left_join(dataset14, PGS_rs, by = "ergoid")
PC_RS1 <- read.table("RS1_PCs.mds", header = T)
PC_RS2 <- read.table("RS2_PCs.mds", header = T)
PC_RS3 <- read.table("RS3_PCs.mds", header = T)
PC_RS <- rbind(PC_RS1, PC_RS2)
PC_RS <- rbind(PC_RS, PC_RS3)
PC_RS <- PC_RS %>%
  rename(ergoid = IID) %>%
  select(ergoid,C1,C2,C3,C4,C5,C6,C7,C8,C9,C10)

dataset4 <- left_join(dataset4, PC_RS, by = "ergoid")

dataset12 <- left_join(dataset11, PC_RS, by = "ergoid")
dataset12 <- left_join(dataset12, PGS_rs, by = "ergoid")

dataset_useProxy <- left_join(dataset12, PRS_RS_USEPROXY, by = "ergoid")
dataset12 <- left_join(dataset12, PGS_77_rs, by = "ergoid")

# generate sex- and age- stratified datasets
## sex
dataset12_f <- subset(dataset12, sex == 1)
dataset12_m <- subset(dataset12, sex == 0)
## age<60yrs and >=60yrs
dataset12_young <- subset(dataset12, age < 60)
dataset12_old <- subset(dataset12, age >= 60)

summary(dataset4$PGS)
hist_PGS <- ggplot(dataset4, aes(x = PGS)) + 
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Histogram of alcohol consumption PGS in the Rotterdam study", 
       x = "Alcohol consumption PGS", 
       y = "Frequency") +
  theme_minimal()
hist_PGS
#ggsave("hist_PGS.png", plot = hist_PGS, width = 6, height = 5)
```

## 1-sample MR processing
```{r}
# 1-sample MR analysis
#library AER package
library(AER)

## LSBMD
iv_ls <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_ls, conf.int = T)

iv_ls_77 <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_77 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_ls_77, conf.int = T)

## LSBMD sex-specific
iv_ls_f <- ivreg(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_f)

iv_ls_m <- ivreg(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_m)

tidy(iv_ls_f, conf.int = T)
tidy(iv_ls_m, conf.int = T)

## LSBMD age-stratified
iv_ls_young <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_young)

iv_ls_old <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_old)

tidy(iv_ls_young, conf.int = T)
tidy(iv_ls_old, conf.int = T)

# LSBMD use proxy
iv_ls_proxy <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_useProxy + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_useProxy)
tidy(iv_ls_proxy, conf.int = TRUE)
```

```{r}
## NKBMD
iv_nk <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_nk, conf.int = T)

iv_nk_77 <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_77 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_nk_77, conf.int = T)

## NKBMD sex-stratified
iv_nk_f <- ivreg(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_f)

iv_nk_m <- ivreg(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_m)

tidy(iv_nk_f, conf.int = T)
tidy(iv_nk_m, conf.int = T)

## NKBMD age-stratified
iv_nk_young <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_young)
iv_nk_old <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_old)

tidy(iv_nk_young, conf.int = T)
tidy(iv_nk_old, conf.int = T)

## NKBMD use proxy
iv_nk_proxy <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_useProxy + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_useProxy)
tidy(iv_nk_proxy, conf.int = T)
```


```{r}
## TOBMD
iv_TOT <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_TOT, conf.int = T)

iv_TOT_77 <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_77 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_TOT_77, conf.int = T)

## TOBMD sex-stratified
iv_TOT_f <- ivreg(std_totbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_f)

iv_TOT_m <- ivreg(std_totbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_m)

tidy(iv_TOT_f, conf.int = T)
tidy(iv_TOT_m, conf.int = T)

## TOBMD age-stratified
iv_TOT_young <- ivreg(std_totbmd ~ log_alc_dag + age + sex + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_young)
iv_TOT_old <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_old)

tidy(iv_TOT_young, conf.int = T)
tidy(iv_TOT_old, conf.int = T)

## TOBMD use proxy
iv_TOT_proxy <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_useProxy + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_useProxy)
tidy(iv_TOT_proxy, conf.int = T)
```

```{r}
## LSTBS
iv_tbs <- ivreg(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_tbs, conf.int = T)

iv_tbs_77 <- ivreg(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_77 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12)
tidy(iv_tbs_77, conf.int = T)

## LSTBS sex-stratified
iv_tbs_f <- ivreg(lstbs ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_f)

iv_tbs_m <- ivreg(lstbs ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_m)

tidy(iv_tbs_f, conf.int = T)
tidy(iv_tbs_m, conf.int = T)

# LSTBS age-stratified
iv_tbs_young <- ivreg(lstbs ~ log_alc_dag + age + sex + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_young)

iv_tbs_old <- ivreg(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset12_old)

tidy(iv_tbs_young, conf.int = T)
tidy(iv_tbs_old, conf.int = T)

## LSTBS use proxy
iv_tbs_proxy <- ivreg(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_useProxy + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_useProxy)
tidy(iv_tbs_proxy, conf.int = T)
```
# Alcohol denpendance SNPs
```{r}
# input genotype data
AD_SNPs_RS1 <- read.table("output_RS1.txt", header = T)
AD_SNPs_RS2 <- read.table("output_RS2.txt", header = T)
AD_SNPs_RS3 <- read.table("output_RS3.txt", header = T)

str(AD_SNPs_RS1)
str(AD_SNPs_RS2)
str(AD_SNPs_RS3)
```

```{r}
# recode genotype into 0/1/2

AD_SNPs_RS1 <- AD_SNPs_RS1 %>%
  mutate(GT1 = case_when(
    GT == "0|0" ~ 2,
    GT == "0|1" ~ 1,
    GT == "1|0" ~ 1,
    GT == "1|1" ~ 0)) %>%
  mutate(GT2 = case_when(
    POS == 100229017 & GT1 == 0 ~ 2,
    POS == 100229017 & GT1 == 2 ~ 0,
    TRUE ~ GT1
  ))

AD_SNPs_RS2 <- AD_SNPs_RS2 %>%
  mutate(GT1 = case_when(
    GT == "0|0" ~ 2,
    GT == "0|1" ~ 1,
    GT == "1|0" ~ 1,
    GT == "1|1" ~ 0)) %>%
  mutate(GT2 = case_when(
    POS == 100229017 & GT1 == 0 ~ 2,
    POS == 100229017 & GT1 == 2 ~ 0,
    TRUE ~ GT1
  ))

AD_SNPs_RS3 <- AD_SNPs_RS3 %>%
  mutate(GT1 = case_when(
    GT == "0|0" ~ 2,
    GT == "0|1" ~ 1,
    GT == "1|0" ~ 1,
    GT == "1|1" ~ 0)) %>%
  mutate(GT2 = case_when(
    POS == 100229017 & GT1 == 0 ~ 2,
    POS == 100229017 & GT1 == 2 ~ 0,
    TRUE ~ GT1
  ))

table(AD_SNPs_RS1$GT2)
table(AD_SNPs_RS2$GT2)
table(AD_SNPs_RS3$GT2)

# modify dataset from long format to wide format
AD_SNPs_RS <- rbind(AD_SNPs_RS1,AD_SNPs_RS2,AD_SNPs_RS3)
AD_SNPs_RS0 <- AD_SNPs_RS[,c(1,4,9)]
AD_SNPs_RS_wide <- AD_SNPs_RS0 %>%
  pivot_wider(names_from = ID, values_from = GT2)
print(AD_SNPs_RS_wide)
```



```{r}
dataset13 <- left_join(dataset12,AD_SNPs_RS_wide,by = "ergoid")
summary(dataset13$`4:100229017`)
summary(dataset13$`4:100239319`)
summary(dataset13$`4:100260789`)
summary(dataset13$`4:100252560`)

dataset14 <- dataset13 %>%
  rename(rs2066702 = `4:100229017`,
         rs1229984 = `4:100239319`,
         rs698 = `4:100260789`,
         rs3114045 = `4:100252560`)

```


# One-sample MR using alcohol dependance SNPs
```{r}
## LSBMD
iv_ls_AD <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | rs2066702 + rs1229984 + rs698 + rs3114045 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset14)
summary(iv_ls_AD)
tidy(iv_ls_AD, conf.int = TRUE)
```

```{r}
## NKBMD
iv_nk_AD <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | rs2066702 + rs1229984 + rs698 + rs3114045 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset14)
summary(iv_nk_AD)
tidy(iv_nk_AD, conf.int = TRUE)
```
```{r}
## TOBMD
iv_to_AD <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | rs2066702 + rs1229984 + rs698 + rs3114045 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset14)
tidy(iv_to_AD, conf.int = TRUE)
```


# Sex-stratified
```{r}
dataset14_f <- subset(dataset14, sex == 1)
dataset14_m <- subset(dataset14, sex == 0)
```

## Female
```{r}
## LSBMD
iv_ls_AD_f <- ivreg(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | rs2066702 + rs1229984 + rs698 + rs3114045 + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset14_f)
summary(iv_ls_AD_f)
```

```{r}
## NKBMD
iv_nk_AD_f <- ivreg(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | rs2066702 + rs1229984 + rs698 + rs3114045 + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset14_f)
summary(iv_nk_AD_f)
```
# Male
```{r}
## LSBMD
iv_ls_AD_m <- ivreg(std_lsbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | rs2066702 + rs1229984 + rs698 + rs3114045 + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset14_m)
summary(iv_ls_AD_m)
```

```{r}
## NKBMD
iv_nk_AD_m <- ivreg(std_nkbmd ~ log_alc_dag + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | rs2066702 + rs1229984 + rs698 + rs3114045 + age + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset14_m)
summary(iv_nk_AD_m)
```

# Combine forest plots of RS and UKB together
```{r}
fp_rs_ukb_ls_fn_tb <-  read_excel("//store/department/im/share/projects/Fernando_Group/Alcohol_BMD_Xinpeng/results/Results/Forestplot_RS_and_UKBB_LS_FN_TO.xlsx")
```

```{r}
fp_rs_ukb_ls_fn_tb |>
  group_by(group) |>
  forestplot(fn.ci_norm = c(fpDrawNormalCI, fpDrawCircleCI),
             boxsize = .25,
             line.margin = .1,
             lineheight = "lines",
             xlab = "Beta (95%CI)") |>
  fp_set_style(box = c("blue", "darkred") |> lapply(function(x) gpar(fill = x, col = "#555555")),
               default = gpar(vertices = TRUE)) |>
  fp_set_zebra_style("#F5F9F9")


```
# Input PAU-PRS
```{r}
# PRS for PAU - 22 SNPs available in both cohorts
PRS_PAU_30_RS1 <- read.table("PGS-SNPs_PAU_30.csv-RS1.txt", header = TRUE)
PRS_PAU_30_RS2 <- read.table("PGS-SNPs_PAU_30.csv-RS2.txt", header = TRUE)
PRS_PAU_30_RS3 <- read.table("PGS-SNPs_PAU_30.csv-RS3.txt", header = TRUE)

PRS_PAU_30_RS <- bind_rows(list(PRS_PAU_30_RS1, PRS_PAU_30_RS2, PRS_PAU_30_RS3))
PRS_PAU_30_RS <- PRS_PAU_30_RS %>%
  rownames_to_column(var = "ergoid") %>%
  rename(PGS_PAU = PGS,
         wPGS_PAU = wPGS)
PRS_PAU_30_RS$ergoid <- as.numeric(PRS_PAU_30_RS$ergoid)

dataset_with_PAUPRS <- left_join(dataset12, PRS_PAU_30_RS, by = "ergoid")

# PRS for PAU - conditionally independent and available 
PRS_PAU_18_RS1 <- read.table("PGS-SNPs_PAU_18.csv-RS1.txt", header = TRUE)
PRS_PAU_18_RS2 <- read.table("PGS-SNPs_PAU_18.csv-RS2.txt", header = TRUE)
PRS_PAU_18_RS3 <- read.table("PGS-SNPs_PAU_18.csv-RS3.txt", header = TRUE)

PRS_PAU_18_RS <- bind_rows(list(PRS_PAU_18_RS1, PRS_PAU_18_RS2, PRS_PAU_18_RS3))
PRS_PAU_18_RS <- PRS_PAU_18_RS %>%
  rownames_to_column(var = "ergoid") %>%
  rename(PGS_PAU_18 = PGS,
         wPGS_PAU_18 = wPGS)
PRS_PAU_18_RS$ergoid <- as.numeric(PRS_PAU_18_RS$ergoid)

dataset_with_PAUPRS <- left_join(dataset_with_PAUPRS, PRS_PAU_18_RS, by = "ergoid")
```

# One-sample MR - PAU
```{r}
## LSBMD
iv_ls_PAU <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_ls_PAU, conf.int = TRUE)
```

```{r}
## NKBMD
iv_nk_PAU <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_nk_PAU, conf.int = TRUE)
```

```{r}
## TOBMD
iv_TOT_PAU <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_TOT_PAU, conf.int = TRUE)
```

```{r}
## LSTBS
iv_tbs_PAU <- ivreg(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_tbs_PAU, conf.int = TRUE)
```

# One-sample MR - PAU - 18
```{r}
## LSBMD
iv_ls_PAU <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU_18 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_ls_PAU, conf.int = TRUE)
```

```{r}
## NKBMD
iv_nk_PAU <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU_18 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_nk_PAU, conf.int = TRUE)
```

```{r}
## TOBMD
iv_TOT_PAU <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU_18 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_TOT_PAU, conf.int = TRUE)
```

```{r}
## LSTBS
iv_tbs_PAU <- ivreg(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_PAU_18 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_with_PAUPRS)
tidy(iv_tbs_PAU, conf.int = TRUE)
```

# One-sample MR SNPs without UKB ----
```{r}
PRS_withoutUKB_RS1 <- read.table("PGS-SNPs_ALC_withoutUKB_RSformat.csv-RS1withoutUKB.txt", header = TRUE)
PRS_withoutUKB_RS2 <- read.table("PGS-SNPs_ALC_withoutUKB_RSformat.csv-RS2withoutUKB.txt", header = TRUE)
PRS_withoutUKB_RS3 <- read.table("PGS-SNPs_ALC_withoutUKB_RSformat.csv-RS3withoutUKB.txt", header = TRUE)

PRS_withoutUKB_RS <- bind_rows(list(PRS_withoutUKB_RS1, PRS_withoutUKB_RS2, PRS_withoutUKB_RS3))
PRS_withoutUKB_RS <- PRS_withoutUKB_RS %>%
  rownames_to_column(var = "ergoid") %>%
  rename(PGS_withoutUKB = PGS,
         wPGS_withoutUKB = wPGS)
PRS_withoutUKB_RS$ergoid <- as.numeric(PRS_withoutUKB_RS$ergoid)

PRS_withoutUKB_RS1_77 <- read.table("PGS-SNPs_ALC_withoutUKB_RSformat_77.csv-RS1withoutUKB_77.txt", header = TRUE)
PRS_withoutUKB_RS2_77 <- read.table("PGS-SNPs_ALC_withoutUKB_RSformat_77.csv-RS2withoutUKB_77.txt", header = TRUE)
PRS_withoutUKB_RS3_77 <- read.table("PGS-SNPs_ALC_withoutUKB_RSformat_77.csv-RS3withoutUKB_77.txt", header = TRUE)

PRS_withoutUKB_RS_77 <- bind_rows(list(PRS_withoutUKB_RS1_77, PRS_withoutUKB_RS2_77, PRS_withoutUKB_RS3_77))
PRS_withoutUKB_RS_77 <- PRS_withoutUKB_RS_77 %>%
  rownames_to_column(var = "ergoid") %>%
  rename(PGS_withoutUKB_77 = PGS,
         wPGS_withoutUKB_77 = wPGS)
PRS_withoutUKB_RS_77$ergoid <- as.numeric(PRS_withoutUKB_RS_77$ergoid)

dataset_PRS_withoutUKB <- left_join(dataset12, PRS_withoutUKB_RS, by = "ergoid")
dataset_PRS_withoutUKB <- left_join(dataset_PRS_withoutUKB, PRS_withoutUKB_RS_77, by = "ergoid")
```

# One-sample MR without UKB
```{r}
## LSBMD
iv_ls_withoutUKB <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_ls_withoutUKB, conf.int = TRUE)
```
```{r}
iv_ls_withoutUKB_77 <- ivreg(std_lsbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB_77 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_ls_withoutUKB_77, conf.int = TRUE)
```


```{r}
## NKBMD
iv_nk_withoutUKB <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_nk_withoutUKB, conf.int = TRUE)
```
```{r}
iv_nk_withoutUKB_77 <- ivreg(std_nkbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB_77 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_nk_withoutUKB_77, conf.int = TRUE)
```


```{r}
## TOBMD
iv_TOT_withoutUKB <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_TOT_withoutUKB, conf.int = TRUE)
```
```{r}
iv_TOT_withoutUKB_77 <- ivreg(std_totbmd ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB_77 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_TOT_withoutUKB_77, conf.int = TRUE)
```


```{r}
## LSTBS
iv_tbs_withoutUKB <- ivreg(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_tbs_withoutUKB, conf.int = TRUE)
```
```{r}
iv_tbs_withoutUKB_77 <- ivreg(lstbs ~ log_alc_dag + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10 | wPGS_withoutUKB_77 + age + sex + rs_cohort.x + C1 + C2 + C3 + C4 + C5 + C6 + C7 + C8 + C9 + C10, data = dataset_PRS_withoutUKB)
tidy(iv_tbs_withoutUKB_77, conf.int = TRUE)
```

